### Statistical Look at each encounter for each recorder type
```{r}
pcplot<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/BIG ROCCA FILE.csv')
```


## Mean of the Mean Positive Slope for MHI
```{r}
df_mhi = droplevels(subset(pcplot, pcplot$population=='mhi'))
posslope_mhi = mean(df_mhi$FREQPOSSLOPEMEAN)
plot(as.factor(df_mhi$group), df_mhi$FREQPOSSLOPEMEAN, xlab="Encounter ID", ylab="Mean Positive Slope", main="Main Hawaiian Island Population")
plot(as.factor(df_mhi$group), df_mhi$DURATION, xlab="Encounter ID", ylab="Mean Positive Slope", main="Main Hawaiian Island Population")
plot(as.factor(df_mhi$group), df_mhi$FREQSWEEPFLATPERCENT, xlab="Encounter ID", ylab="% of the whistle with zero slope ", main="Main Hawaiian Island Population")

posslope_mhi = mean(df_mhi$FREQNEGSLOPEMEAN)
plot(df_mhi$id, df_mhi$FREQNEGSLOPEMEAN, xlab="Encounter ID", ylab="Mean Negative Slope", main="Main Hawaiian Island Population")

plot(df_mhi$id, df_mhi$DURATION, xlab="Encounter ID", ylab="Mean Negative Slope", main="Main Hawaiian Island Population")
plot(df_mhi$id, df_mhi$FREQSWEEPFLATPERCENT, xlab="Encounter ID", ylab="Mean Negative Slope", main="Main Hawaiian Island Population")
plot(df_mhi$id, df_mhi$FREQSLOPEMEAN, xlab="Encounter ID", ylab="Mean Negative Slope", main="Main Hawaiian Island Population")


mhi30 = filter(df_mhi, df_mhi$group == 30)
#Mean of mean positive slope for each  mhi encounter 
write.csv(list(aggregate(df_mhi$FREQPOSSLOPEMEAN, by=list(df_mhi$id), FUN=mean)),'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\MHI_MeanPosSlope.csv') 
aggregate(df_mhi$FREQNEGSLOPEMEAN, by=list(df_mhi$id), FUN=mean)
aggregate(df_mhi$FREQPOSSLOPEMEAN, by=list(df_mhi$Recorder), FUN=median)
aggregate(df_mhi$DURATION, by=list(df_mhi$Recorder), FUN=median)
```

mean(mhi30redo$FREQPOSSLOPEMEAN)

library(lattice)
## Mean of the Mean Positive Slope for PELAGIC
df_pel = droplevels(subset(pcplot, pcplot$population=='pelagic'))
posslope_pel = mean(df_pel$FREQPOSSLOPEMEAN)
plot(as.factor(df_pel$group), df_pel$FREQPOSSLOPEMEAN, xlab="Encounter ID", ylab="Mean Positive Slope", main= "Pelagic Population")

posslope_pel = mean(df_pel$FREQNEGSLOPEMEAN)
plot(as.factor(df_pel$group), df_pel$FREQNEGSLOPEMEAN, xlab="Encounter ID", ylab="Mean Negative Slope", main= "Pelagic Population")

plot(as.factor(df_pel$group), df_pel$FREQSWEEPFLATPERCENT, xlab="Encounter ID", ylab="% of the whistle with zero slope ", main="Pelagic Population")
plot(as.factor(df_pel$group), df_pel$FREQQUARTER3, xlab="Encounter ID", ylab="% of the whistle with zero slope ", main="Pelagic Population")

write.csv(list(aggregate(df_pel$FREQPOSSLOPEMEAN, by=list(df_pel$id), FUN=mean)),'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\Pelagic_MeanPosSlope.csv') 
aggregate(df_pel$FREQNEGSLOPEMEAN, by=list(df_pel$id), FUN=mean)
aggregate(df_pel$FREQPOSSLOPEMEAN, by=list(df_pel$Recorder), FUN=median)
aggregate(df_pel$FREQQUARTER3, by=list(df_pel$id), FUN=mean)


## Mean of the Mean Positive Slope for NWHI
df_nwhi = droplevels(subset(pcplot, pcplot$population=='nwhi'))
posslope_nwhi = mean(df_nwhi$FREQPOSSLOPEMEAN)
plot(as.factor(df_nwhi$group), df_nwhi$FREQPOSSLOPEMEAN, xlab="Encounter ID", ylab="Mean Positive Slope", main="Northwestern Hawaiian Island Population")

posslope_nwhi = mean(df_nwhi$FREQNEGSLOPEMEAN)
plot(df_nwhi$id, df_nwhi$FREQNEGSLOPEMEAN, xlab="Encounter ID", ylab="Mean Negative Slope", main="Northwestern Hawaiian Island Population")


plot(as.factor(df_nwhi$group), df_nwhi$FREQSWEEPFLATPERCENT, xlab="Encounter ID", ylab="% of the whistle with zero slope ", main="Northwestern Hawaiian Island Population")

write.csv(list(aggregate(df_nwhi$FREQPOSSLOPEMEAN, by=list(df_nwhi$id), FUN=mean)),'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\NWHI_MeanPosSlope.csv')
aggregate(df_nwhi$FREQPOSSLOPEMEAN, by=list(df_nwhi$id), FUN=mean)
aggregate(df_nwhi$FREQNEGSLOPEMEAN, by=list(df_nwhi$id), FUN=mean)
aggregate(df_nwhi$FREQPOSSLOPEMEAN, by=list(df_nwhi$Recorder), FUN=median)
aggregate(df_nwhi$FREQQUARTER3, by=list(df_nwhi$id), FUN=mean)




plot(as.factor(pcplot$group), pcplot$FREQQUARTER3, col=pcplot$population, xlab="Encounter ID", ylab="Frequency of the third quarter of the whistle")

```{r}
library(ggplot2)
#shades of gray###
color <- c('#35bb7c', 'black', '#947d51','#8576aa')


#Plot Mean positive slope / duration
freqslope=ggplot(pcplot, aes(x = EncounterID, y = FREQPOSSLOPEMEAN, col=as.factor(population))) +
  geom_boxplot() +
  theme_bw() +
  labs(x="Encounter", y="Frequency (Hz)", 
       title="Mean Positive Slope") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=7),
        axis.text.y = element_text(size=9),
        axis.title=element_text(size=12,face="bold"),
        legend.title = element_blank()) +
  scale_color_manual(values=color) 
ggsave('Plot_FREQPOSSLOPE.pdf', width = 11, height = 8, units = "in")




# Plot third-quarter frequency / duration
freqspd<-ggplot(pcplot, aes(x = EncounterID, y = FREQSPREAD, col=as.factor(population))) +
  geom_boxplot() +
  theme_bw() +
  labs(x="Encounter", y="Frequency (Hz)")+ 
       #title="Third Quarter Frequency", 
       #subtitle="Frequency at three-quarters of the duration") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=7),
        axis.text.y = element_text(size=9),
        axis.title=element_text(size=12,face="bold"),
        legend.title = element_blank()) +
  scale_color_manual(values=color) 
 
  

ggsave('Plot_freqquarter3.png', width = 11, height = 8, units = "in", type = "cairo-png")




# Plot FREQSWEEPFLATPERCENT
freqswp <- ggplot(pcplot, aes(x = EncounterID, y = pcplot$FREQSWEEPFLATPERCENT, col=as.factor(population))) +
  geom_boxplot() +
  theme_bw() +
  labs(x="Encounter", y="", 
       title="Percent Flat",
       subtitle="Percent of the whistle with zero slope") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=7),
        axis.text.y = element_text(size=9),
        axis.title=element_text(size=12,face="bold"),
        legend.title = element_blank()) +
  guides(shape = guide_legend(override.aes = list(size=10)))+
  scale_color_manual(values=color)
ggsave('Plot_FREQSWEEPFLAT.pdf', width = 11, height = 8, units = "in")
#ggsave('Plot_FREQSWEEPFLAT.png', device='pdf', width = 11, height = 8, units = "in")
```

```{r}
# Plot Accuracy by Duration
pcACC <- read.csv("C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data/Barkley_Sensitivity Analysis Table.csv")

#purple, green, orange
colorPOP <- c('#791fb8','#2b961a','darkorange2')

ggplot(pcACC, aes(x = pcACC$duration..s., y = accuracy, col=as.factor(Recorder))) +
  
  theme_bw() +
  labs(x="Duration (s)", y="Percent Accuracy")+ 
  geom_text(aes(label=pcACC$EncounterPoster), size=5, vjust=-0.1)+
  geom_point() +
  #title="Percent Flat", 
  #subtitle="Percent of the whistle with zero slope") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12),
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_blank(),
        legend.text=element_text(size=15)) +
  guides(shape=guide_legend(override.aes = list(shape = 16)))+
  scale_color_manual(values=color) 
ggsave('Plot_AccVSDur.pdf', width = 11, height = 8, units = "in")


#just to get legend with dots

ggplot(pcACC, aes(x = pcACC$duration..s., y = accuracy, col=as.factor(Recorder))) +
  
  theme_bw() +
  labs(x="Duration (s)", y="Percent Accuracy")+ 
  geom_point() +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12),
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_blank(),
        legend.text=element_text(size=15)) +
  guides(color = guide_legend(override.aes = list(size=5)))+
  scale_color_manual(values=color)
ggsave('Plot_AccVSDurLEGEND.pdf', width = 11, height = 8, units = "in")


ggplot(pcACC, aes(x = pcACC$Recorder, y = accuracy, col=as.factor(Population))) +
  
  theme_bw() +
  labs(x="Recorder Type", y="Percent Accuracy")+ 
  geom_point()+
  geom_text(aes(label=pcACC$EncounterPoster), size=4, hjust=0, nudge_x=0.05)+
  #title="Percent Flat", 
  #subtitle="Percent of the whistle with zero slope") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=12, face='bold'),
        axis.text.y = element_text(size=12, face='bold'),
        axis.title=element_text(size=16,face="bold"),
        legend.position = "none") + #no legend
  guides(shape=guide_legend(override.aes = list(shape = 16)))+
  scale_color_manual(values=colorPOP) 
ggsave('Plot_AccVSDur.png', width = 11, height = 8, units = "in", type = "cairo-png")
```

```{r}
#Explore Cluster 4, MHI 29

mhi29 <- subset(pcplot, pcplot$group == 30)

mhi.only <- subset(pcplot, pcplot$population == 'MHI')
```
