---
title: "Encounter Classifier"
author: "Yvonne Barkley"
date: "April 16, 2018"
output: html_document
---

```{r warning=FALSE, message=FALSE}
library(sqldf)
library(randomForest)
library(party)
library(caret)
library(corrplot)
library(parallel); 
library(doParallel);
library(gtools)
library(devtools) 
library(dplyr)
library(data.table)
library(ggplot2)


```

```{r}
#pcdata <- read.csv('C:\\Users\\yvonne.barkley\\Documents\\PHD\\Chp1_PcClassifier-master\\BIG ROCCA FILE.csv')
pcdata <-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/BIG ROCCA FILE.csv')
pcdata <- droplevels(filter(pcdata, population != "nwhi"))

PEL = 1:8                # designated groupID's assigned for each pelagic group 
#NWHI  = 14:17   # designated groupID's assigned for each NWHI group, #20 was absorbed into #19 (June 2017)     
MHI = c(26:30)    # designated groupID's assigned for each MHI group, #25 absorbed into #26 (June 2017) 
                              # Oct 7: more combining of NWHI and MHI occurred. Performed in Pc_DatManipulation.rmd
#This is set up to be flexible to change later if needed
nrep  = 5           # multiplier for total number of groups
npel  = 1           # number of pelagic schools to group together
nnwhi = 1           # number of northwest HI schools to group together
nmhi  = 1           # number of main HI schools to group together
w = 90              # Number of whistles randomly pulled from each group 
ntest = 1           # number of groups to randomly pull for test dataset
n_samps = 10        #Zack: number of times you want to resample from your dataset
prop = 0.5
```


```{r}



ptm <- proc.time()

results.enc = matrix(0, nrow = 13, ncol = 13)
realP = c()
fakeP = c()
realM = c()
fakeM = c()
output_WhistleClass.enc =NULL
output_EncounterClass.enc=NULL
output_RawResults.enc=NULL
output_TrainSet.encAll=NULL
output_TestSet.encAll = NULL
output_Votes.encAll=NULL
tuned.encAll=NULL
varIMPtotal.enc =NULL
## GIANT LOOP---BEWARE ####

falsehope=c()

for(rep in 1:n_samps){


#From all of the encounters, sampling 50 whistles from each encounter (some have >50)
enc_all = NULL
for(i in c(PEL,MHI)){
  enc_sub=droplevels(subset(pcdata, pcdata$group==i)) #selects all whistles from the randomly selected groups, drops empty levels too
  enc_samp=enc_sub[sample(nrow(enc_sub), w),] #randomly samples w whistles from the 50 or more whistles
  enc_all <-rbind(enc_all, enc_samp)
  
}

enc_part <- createDataPartition(y=enc_all$EncounterID,p=prop,list=FALSE)
enc_train <- enc_all[enc_part,]
enc_test <- enc_all[-enc_part,]



##############################
## Set up a dataframe where each row is a specific 
## combination of mtry (# of variables selected at each split), n_tree (# of trees in  forest), possibly other stuff
##############################
mtry_vals = 5:7 
ntree_vals = seq(501,3001,250) #from, to, by
tune_settings2 = expand.grid(mtry = mtry_vals, ntree = ntree_vals)

#################################
## Initiate Cluster
#################################
cl = makeCluster(detectCores())
registerDoParallel(cl, cores = detectCores())

############################################
##Create the formula for the RF model with 'population' as the response and time-frequency measurements from uncorrelated data as the covariates
############################################
#formPc.new <- as.formula(paste("population ~ ", paste(names(pcnew)[7:36],collapse="+")))
formPc_enc <- as.formula(paste("EncounterID ~ ", paste(names(pcdata)[7:54],collapse="+"))) #Use variables to classify to encounter

#######################################
##The foreach function is the loop function for parallel processing. 
##It uses %dopar% which partitions the loop among your cores, then combines each output using the .combine argument. 

x = foreach(i = 1:nrow(tune_settings2), .combine = rbind) %dopar% { 
  library(randomForest)
  do_RF2 = randomForest(formPc_enc, data = enc_train, test = enc_test, ntree = tune_settings2$ntree[i], mtry = tune_settings2$mtry[i])
  print(do_RF2$err.rate[tune_settings2$ntree[i],])
}

#so, for each of the parameter combinations, i.e., rows of the tune_settings df, the RF is run using those 
#settings (line 46) and the error rates for the 3 pop's and the oob are printed (line 47). 
#when all the rows are finished, the results are combined using the rbind function into one matrix, stored as x.

#NB: .combine can be anything, really, for example .combine = c for a vector, .combine = list for a list, etc.

#################################
## Stop Cluster
#################################
stopCluster(cl)

########################################
## Combine settings and results
########################################
tune_settings2 = cbind(tune_settings2, x)
########################################
## Optimal Settings: the settings combo
## that has the lowest OOB rate
#######################################
tuned.enc<-tune_settings2[which.min(tune_settings2$OOB),]
tuned.encAll <- rbind(tuned.encAll, tuned.enc)
 
### RANDOM FOREST ####
accuracy.enc <- c()
enc_pred.table=NULL


##Run Random Forest model using 'tuned' parameters for mtry and ntree
enc_fit <- randomForest(formPc_enc, data=enc_train, ntree=tuned.enc[1,2], importance=TRUE, replace=T, mtry=tuned.enc[1,1], norm.votes=T, confusion=T, keep.inbag=T, proximity=T) 

enc_prediction <-predict(enc_fit, newdata=enc_test, predict.all = F, type='response') #prediction for encounter of test data using RF model, 'fit'
enc_prediction_vote <-predict(enc_fit, newdata=enc_test, predict.all = T, type='vote') #prediction for encounter of test data using RF model, 'fit'
enc_test$rightPred <- as.character(enc_prediction) == as.character(enc_test$EncounterID) #shows T/F for individual whistles and whether they were classified correctly
accuracy.enc <- sum((enc_test$rightPred)/nrow(enc_test)) #sums all 'TRUE' and divides by total

###RESULTS!!!##################################  
enc_results.temp <- table(enc_test$EncounterID, enc_prediction)    


# Important Variables
  varIMPtemp.enc <- as.data.frame(importance(enc_fit))
  varIMP.enc <- varIMPtemp.enc[order(varIMPtemp.enc$MeanDecreaseGini), , drop =F]
  varIMPtotal.enc <- rbind(varIMPtotal.enc, "REP"=rep, varIMP.enc)


results.enc <- results.enc + enc_results.temp  


}

proc.time()-ptm
```

#enc_pred.table<-rbind(enc_pred.table, rep, enc_results, accuracy2=signif(accuracy2, digits = 3)) #rough table keeping track of confusion matrices for all iterations

write.table(results.enc, 'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/Encounters.csv', sep=",", append=TRUE)
            
enc.whistles<-data.frame("group"=enc_test$EncounterID, "prediction"=enc_prediction, "whistle"=enc_test$whistle_count)

false=nrow(subset(enc_test, enc_test$rightPred=='FALSE'))/nrow(enc_test)*100 #percentage of misclassified whistles
falsehope <- c(falsehope, false)

## OUTPUTS ####
#   d.enc = as.data.frame(unique(enc_test$id))
#   d.enc = as.data.frame(as.character(d.enc[27, ]))
#   e.enc=c()
#   e.enc=as.data.frame(rbind(true.pel.array))#, true.nwhi.array))
#   e.enc=cbind(e.enc, d.enc)
#   colnames(e.enc) = c("Classification", "EncounterID")
#   
# f.array=as.data.frame(cbind(prediction.array.vote$aggregate, prediction.array.vote$individual))
# g.array=cbind(enc, "population"=array.test$population, "group"=array.test$group, whistle=array.test$whistle_count, "prediction"=array.test$rightPred, f.array)
# 
# output_VotesAll.array <- bind_rows(output_VotesAll.array, g.array)


}
```
