---
title: "False Killer Whale Whistle Classification- Random Forest Analysis"
author: "Yvonne Barkley"
date: "March 21, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

This script creates random forests using whistle datasets (either old or new datasets)



# Load your libraries
```{r warning=FALSE, message=FALSE}
#library(sqldf)
library(randomForest)
library(partykit)
library(caret)
library(corrplot)
library(parallel); 
library(doParallel);
library(gtools)
library(devtools) 
library(dplyr)
#library(ROCR)
library(Hmisc)
library(tree)
```



If using the raw data file, run this section to modify the columns
```{r}
#6/28/18
#Load ROCCA data (already replaced 'NA' and '999' with '0')
ROCCAdata<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/ROCCA Towed Data/BigRoccaFile_PM_TowedAll_RAW_20180717.csv')

#ROCCAdata<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/decimation_test/PcMHI31_32_RoccaContourStats_RAWDEC.csv')

#Read old dataset
#pcdata<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\Pc_MasterListEdit-1016labs.csv')

#pcdata<-droplevels(subset(pcdata, Recorder != 'PMRF'))


#Add species names to last columns and save (used this for BIG ROCCA with all recorders)
colnames(ROCCAdata)[93] <- "Dc_Dd"
colnames(ROCCAdata)[94] <- "Gm"
colnames(ROCCAdata)[95] <- "Pc"
colnames(ROCCAdata)[96] <- "Sa"
colnames(ROCCAdata)[97] <- "Sb"
colnames(ROCCAdata)[98] <- "Sc"
colnames(ROCCAdata)[99] <- "Sl"
colnames(ROCCAdata)[100] <- "Tt"

#Used this routine for towed only data set
# colnames(ROCCAdata)[92]
# colnames(ROCCAdata)[93]
# colnames(ROCCAdata)[94]
# colnames(ROCCAdata)[95]
# colnames(ROCCAdata)[97] <- "Dc_Dd"
# colnames(ROCCAdata)[98] <- "Gm"
# colnames(ROCCAdata)[99] <- "Pc"
# colnames(ROCCAdata)[100] <- "Sa"
# colnames(ROCCAdata)[101]<- "Sb"
# colnames(ROCCAdata)[102]<- "Sc"
# colnames(ROCCAdata)[103]<- "Sl"
# colnames(ROCCAdata)[104]<- "Tt"


#Reexamining whistles with outlying values of slope means
# test<- filter(ROCCAdata, FREQPOSSLOPEMEAN  > 90000)
# test2<- filter(ROCCAdata, FREQNEGSLOPEMEAN <  -90000)

#Remove unnecessary columns

pcdata <- ROCCAdata[, -c(1:7, 11:13, 21, 22, 34:37, 68:92, 101)] #for towed only

#for decimation test
#pcdata <- ROCCAdata[, -c(1:7, 11:13, 15:18, 25, 26, 38:41, 73:96, 105)]
pcdata %>% mutate_if(is.factor, as.character) -> pcdata
#not sure why I included the opposite code below...
#pcdata %>% mutate_if(is.character, as.factor) -> pcdata


#i <- sapply(pcdata, is.factor)
#pcdata[i] <- lapply(pcdata[i], as.character)
###Added group column manually
###Add Population column and reorder
pcdata$population <-NA
levels(pcdata$population) <- c(levels(pcdata$population), "pel", "mhi") #, "nwhi")
pcdata[pcdata$group %in% c(31:34), "population"] <- "mhi"
#pcdata[pcdata$group %in% c(14:17), "population"] <- "nwhi"
pcdata[pcdata$group %in% c(1:8), "population"] <- "pel"
pcdata$population <- as.factor(pcdata$population)
pcdata.pm<- pcdata[, c(60, 1:59)]

#Change column namec
colnames(pcdata.pm)[2] <- "EncounterID"

#save file
write.csv(pcdata.pm, 'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/ROCCA Towed Data/BigRoccaFile_PM_TowedAll_EDIT_20180717.csv', row.names=F)

```

Correlation Stuff
Will remove correlated variables based on the r value.
```{r}
##Correlation?#### 
# Including correlated variables would mostly impact the important variables
##Calculate Pearson's correlation coefficient for all variables. Remove variables -0.70 > r > 0.70.
pcsub<-pcdata.pm[,6:52]

#pcsub<-pc_HarpFostex[,7:54]

corrplot.mixed(cor(pcsub, method= "pearson"), upper="number", lower="circle")
tmp <- cor(pcsub)
tmp[upper.tri(tmp)] <- 0
diag(tmp) <- 0
pcsubcor <- pcsub[,!apply(tmp,2,function(x) any(x > 0.80 | x < -0.80))]

#remove FreqCOFM
pcsubcor <-pcsubcor[, -6]
#Adding beginning freq, range, freqquarter3. See Correlation_20180717.xlsx for more details. 
pcsubcor <-cbind(pcsubcor, pcdata.pm[, c(9,11)]) #add in FreqBeg & Range
pcsubcor <-cbind(pcsubcor, pc_HarpFostex[, c(10,12)]) #add in FreqBeg & Range
tmp_new <- cor(pcsubcor)
tmp_new[upper.tri(tmp_new)] <- 0
diag(tmp_new) <- 0
corrPlot<-corrplot.mixed(cor(pcsubcor), upper="circle", lower="number") # for a visual representation of the correlation
write.csv(tmp_new, 'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results/20180717/Correlation_2_20180717.csv', row.names=F)  


#Adding first 5 cols plus a few freq variables that were correlated and removed altogether. 
  pcdata.pm <- cbind(pcdata.pm[, c(1:5)], pcsubcor)
  pc_HarpFostex <- cbind(pc_HarpFostex[, c(1:5)], pcsubcor)

 write.csv(pc_HarpFostex, 'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results/20180720/pcdata.HarpFostex_201807123.csv', row.names=F)   
  


#Example Correlation Code
# https://stats.stackexchange.com/questions/50537/should-one-remove-highly-correlated-variables-before-doing-pca
n.cases <- 240               # Number of points.
n.vars <- 4                  # Number of mutually correlated variables.
set.seed(26)                 # Make these results reproducible.
eps <- rnorm(n.vars, 0, 1/4) # Make "1/4" smaller to *increase* the correlations.
x <- matrix(rnorm(n.cases * (n.vars+2)), nrow=n.cases)
beta <- rbind(c(1,rep(0, n.vars)), c(0,rep(1, n.vars)), cbind(rep(0,n.vars), diag(eps)))
y <- x%*%beta                # The variables.
cor(y)                       # Verify their correlations are as intended.
plot(data.frame(y))          # Show the scatterplot matrix.

# Perform PCA on the first 2, 3, 4, ..., n.vars+1 variables.
p <- lapply(2:dim(beta)[2], function(k) prcomp(y[, 1:k], scale=TRUE))

# Print summaries and display plots.
tmp <- lapply(p, summary)
par(mfrow=c(2,2))
tmp <- lapply(p, plot)


```


#Set up encounters for RF using new or old whistle data
```{r}
#all whistles for pel and mhi only
#pcdata.pm <- read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/ROCCA Towed Data/BigRoccaFile_PMN_TowedAll_EDIT_20180816.csv')

#pcdata.pm <- read.csv('C:\\Users\\yvers\\Documents\\CHP1\\BigRoccaFile_PMN_TowedAll_EDIT_20180816.csv')

pcdata.pm <- read.csv('C:\\Users\\yvers\\Documents\\CHP 1\\data\\ROCCA Towed Data\\BigRoccaFile_PMN_TowedAll_EDIT_20181124.csv')



#try using only top 5 important variables
#pcdata.pm <- pcdata.pm[, c(1:6, 8, 12:15)]

# Need even number of encounters for each population represented in the training data. 
PEL = 1:8                # designated groupID's assigned for each pelagic group 
#NWHI  = 14:17   # designated groupID's assigned for each NWHI group, #20 was absorbed into #19 (June 2017)     
MHI = c(31:34)    # designated groupID's assigned for each MHI group, #25 absorbed into #26 (June 2017) 
                              # Oct 7: more combining of NWHI and MHI occurred. Performed in Pc_DatManipulation.rmd

nrep  = 4           # multiplier for total number of groups
npel  = 1           # number of pelagic schools to group together
nnwhi = 1           # number of northwest HI schools to group together
nmhi  = 1           # number of main HI schools to group together
t = 0.6
w = 150              # Number of whistles randomly pulled from each group 
ntest = 1           # number of groups to randomly pull for test dataset
n_samps = 100         #number of times you want to resample from your dataset


```


Meat of the Analysis

```{r message=FALSE, warning=FALSE}

ptm<-proc.time() 


#### Start Main Analysis ####
results_pm = matrix(0, nrow = 2, ncol = 2)
output_ModelFit.pm = NULL
realP.pm = c()
realM.pm = c()
#use this format of output_WhistleClass on desktop (R version 3.3.1)
output_WhistleClass = NULL
#use this format of output_WhistleClass on laptop (R version 3.5.1)
# output_WhistleClass = data.frame(Doubles=double(),
#                  Ints=integer(),
#                  Factors=factor(),
#                  Logicals=logical(),
#                  Characters=character(),
#                  stringsAsFactors=FALSE)
output_EncounterClass.pm=NULL
output_results_pm=NULL
output_TrainSetAll.pm=NULL
output_TestSetAll.pm = NULL
output_Votes.pmAll=NULL
tunedAll.pm=NULL
output_varIMPtotal.pm = NULL
fit_error_rate.pm = c()
accuracies.pm = c()
pc_classPM = NULL

#### OVERALL LOOP TO BUILD RANDOM FORESTS ####  
for(rep in 1:n_samps){ 
 
  set.seed(rep) #YB placed the seed inside the loop bc it wasn't duplicating trees
  
  
  #This randomly samples 'nrep' groups from each population
  N_pel = as.numeric(sample(PEL, nrep*npel))
  N_mhi  = as.numeric(sample(MHI, nrep*nmhi))
  #N_nwhi  = as.numeric(sample(NWHI, nrep*nnwhi))
 
  
  # Test data 
  #Randomly sample 1 group out of the selected groups for each population to be included in the test dataset later
  test_pel= as.integer(sample(N_pel, ntest))
  test_mhi= as.integer(sample(N_mhi, ntest))

  
  #Combines test groups into single vector
  N_test = rbind(test_pel, test_mhi)
  
  all.pm = NULL
  for(i in c(N_pel,N_mhi)){
    sub=droplevels(subset(pcdata.pm, pcdata.pm$group==i)) #selects all whistles from the randomly selected groups, drops empty levels too
    samp=sub[sample(nrow(sub), w),] #randomly samples w whistles from the 50 or more whistles
    all.pm <-rbind(all.pm, samp)
    
    }
  #Correlation Step####
pcsub.pm<-all.pm[,7:53]
#corrplot.mixed(cor(pcsub, method= "pearson"), upper="number", lower="circle")
tmp.pm <- cor(pcsub.pm)
tmp.pm[upper.tri(tmp.pm)] <- 0
diag(tmp.pm) <- 0
pcsubcor.pm <- pcsub.pm[,!apply(tmp.pm,2,function(x) any(x > 0.80 | x < -0.80))]

all.pm <- cbind(all.pm[, c(1:6)], pcsubcor.pm)
  

test.pm<-droplevels(subset(all.pm, group%in%N_test ))
train.pm<-droplevels(subset(all.pm, !(group%in%N_test)))


#### Tuning parameters ####
## Set up a dataframe where each row is a specific combination of mtry (# of variables selected at each split), n_tree (# of trees in  forest), possibly other stuff
mtry_vals = sqrt(length(all.pm[,7:length(all.pm)]))
ntree_vals = seq(501,5001,500) #from, to, by
tune_settings = expand.grid(mtry = mtry_vals, ntree = ntree_vals)#, nodesize = node_vals)
#tune_settings2 = expand.grid(mtry = mtry_vals, ntree = ntree_vals)

#### Run Optimization Sequence ####
## Initiate Cluster

cl = makeCluster(detectCores())
registerDoParallel(cl, cores = detectCores())


#### Formula ####
# for the RF model with 'population' as the response and time-frequency measurements 

formPc.pm <- as.formula(paste("population ~ ", paste(names(all.pm)[7:length(all.pm)],collapse="+")))


## TOP 5 UNCORRELATED VARIABLES
#formPc <- as.formula(paste("population ~ ", paste(names(pcdata.pm)[7:11],collapse="+")))

## CORRELATED VARIABLES
#formPc <- as.formula(paste("population ~ ", paste(names(pcdata.pm)[7:54],collapse="+"))) #Use variables to classify to population



##The foreach function is the loop function for parallel processing. 
##It uses %dopar% which partitions the loop among your cores, then combines each output using the .combine argument. 

x.pm = foreach(i = 1:nrow(tune_settings), .combine = rbind) %dopar% { 
    library(randomForest)
    do_RF = randomForest(formPc.pm, data = train.pm, test = test.pm, ntree = tune_settings$ntree[i], mtry = tune_settings$mtry[i])#, nodesize = tune_settings$nodesize[i])
    print(do_RF$err.rate[tune_settings$ntree[i],])
  }

#so, for each of the parameter combinations, i.e., rows of the tune_settings df, the RF is run using those 
#settings (line 188) and the error rates for the 3 pop's and the oob are printed (line 189). 
#when all the rows are finished, the results are combined using the rbind function into one matrix, stored as x.

#NB: .combine can be anything, really, for example .combine = c for a vector, .combine = list for a list, etc.


#### Stop Cluster ####

stopCluster(cl)

## Combine settings and results ####

tune_settings.pm = cbind(tune_settings, x.pm)


## Optimal Settings: the settings combo
## that has the lowest OOB rate

tuned.pm<-tune_settings.pm[which.min(tune_settings.pm$OOB),]
tunedAll.pm <- rbind(tunedAll.pm, tuned.pm)

#### Run Random Forest ####
##Run Random Forest model using 'tuned' parameters for mtry and ntree
  fit.pm <- randomForest(formPc.pm, data=train.pm, ntree=tuned.pm[1,2], importance=TRUE, replace=T, mtry=tuned.pm[1,1], norm.votes=F, confusion=T, keep.inbag=T, proximity=T, keep.forest=T, scale = F) 



#prediction for population of test data using RF model, 'fit', returns a list with aggregate and individual votes for each whistle  
  prediction.pm <-predict(fit.pm, newdata=test.pm, predict.all = F, type='response', nodes = T) 
  prediction.vote <-predict(fit.pm, newdata=test.pm, predict.all = T, type='vote')

   #shows T/F for individual whistles and whether they were classified correctly  
   test.pm$rightPred <- as.character(prediction.pm) == as.character(test.pm$population) 
  
   #sums all 'TRUE' and divides by total
   accuracy.pm <- sum((test.pm$rightPred)/nrow(test.pm)) 
   accuracies.pm <- c(accuracies.pm, accuracy.pm)
  
   ####RESULTS!!!#####  
results_pmtemp <- table(test.pm$population, prediction.pm)  # conf matrix of individual whistle counts

#Summed results confusion matrix after each iteration
results_pm = results_pm + results_pmtemp    
   
# 3/23/19   
#### Classify Encounters by Majority Vote ####

pc_classPM_pred <- as.data.frame(colnames(results_pmtemp)[max.col(results_pmtemp)])

pc_classPM_temp <-cbind(rep, unique(test.pm$EncounterID), as.data.frame(row.names(fit.pm$confusion)), pc_classPM_pred)

pc_classPM <- rbind(pc_classPM, pc_classPM_temp)   
   
   
   
   
################### 12/10/18
   #Set a threshold for the percentage of whistles that must be correctly classified to classify the population
# true.pel = if(results_pmtemp[1,1] >= t*w) 'Pelagic' else 'Ambiguous'
# #pel.nwhi = if(results.pntemp[1,2] >= t*w) 'NWHI' else 'Ambiguous'
# pel.mhi = if(results_pmtemp[1,2] >= t*w) 'MHI' else 'Ambiguous'
# 
# #true.nwhi = if(results_pmtemp[2,2] >= t*w) 'NWHI' else 'Ambiguous'
# #nwhi.pel = if(results.pntemp[2,1] >= t*w) 'Pelagic' else 'Ambiguous'
# #nwhi.mhi = if(results_pmtemp[2,3] >= t*w) 'MHI' else 'Ambiguous'
# 
# true.mhi = if(results_pmtemp[2,2] >= t*w) 'MHI' else 'Ambiguous'
# mhi.pel = if(results_pmtemp[2,1] >= t*w) 'Pelagic' else 'Ambiguous'
# #mhi.nwhi = if(results.temp[3,2] >= t*w) 'NWHI' else 'Ambiguous'
# 

##number of whistles correctly classified per population####
true.enc = data.frame(rbind(results_pmtemp[1,1], results_pmtemp[2,2]))
#d = as.data.frame(unique(test$EncounterID))
#d = as.data.frame(d[c(1:3), ])
e.pm=c()
f=c()
#h=c()
e.pm=as.data.frame(rbind(true.pel, pel.mhi))
#f=as.data.frame(rbind(nwhi.pel, true.nwhi))
h=as.data.frame(rbind(mhi.pel, true.mhi))
true.enc = data.frame(rbind(results_pmtemp[1,1], results_pmtemp[2,2]))
e.pm=cbind(as.data.frame(N_test), e.pm, h, true.enc)
colnames(e.pm) = c("TrueEnc", "Pelagic", "MHI", "CorrectDW" ) #, "EncounterID")

   


#For calculating means and standard dev for EACH ITERATION of confusion matrix
#For a single iteration, stores separate cells of class results for calculating mean and standard error
#YOU NEED THIS STEP 4/26/18, TRUST ME!
realP.pm <- c(realP.pm,results_pmtemp[1,1])
#fakeP <- c(fakeP,results_pmtemp[1,2])
realM.pm <- c(realM.pm,results_pmtemp[2,2])
#fakeM <- c(fakeM,results_pmtemp[2,1])
        
#### OUTPUTS ####
#1.
#Save each RF model's confusion matrix and OOB error
output_ModelFit.pm <- rbind(output_ModelFit.pm, as.table(fit.pm$confusion), mean(fit.pm$err.rate))

#2.
#Most comprehensive information about overall predictions, proportional classification of trees, and tree predictions for each whistle
g.pm=cbind(rep, "EncounterID"= test.pm$EncounterID, "EncounterCount"=test.pm$EncounterCount, "Prediction"=test.pm$rightPred, "PredictedPop"=prediction.pm,as.data.frame(cbind(prediction.vote$aggregate, prediction.vote$individual)))

#output_WhistleClass.pm <- bind_rows(list(output_WhistleClass.pm, g.pm), .id = 'id')

#3.
#table keeping track of confusion matrices for all iterations
output_results_pm<-rbind(output_results_pm, rep, results_pmtemp, accuracy.pm=signif(accuracy.pm, digits = 3)) 

#4.
#Output of encounters used in each training set  
output_TrainSet.pm <- data.frame(rep, "EncounterID" = sort(unique(train.pm$EncounterID))) #, "encounter" = sort(unique(train$group)))
output_TrainSetAll.pm <-rbind(output_TrainSetAll.pm, output_TrainSet.pm)

#5.
#Output shows encounters in each test set, encounter prediction, and % correctly classified
output_TestSet.pm <- data.frame(rep, "EncounterID" = sort(unique(test.pm$EncounterID)), e.pm) #, "encounter" = sort(unique(train$group)))
output_TestSetAll.pm <-rbind(output_TestSetAll.pm, output_TestSet.pm)

#6.
  # Important Variables
  varIMPtemp.pm <- as.data.frame(importance(fit.pm))
  varIMP.pm <- varIMPtemp.pm[order(varIMPtemp.pm$MeanDecreaseAccuracy), , drop =F]
  varIMP.pm$Rep <- rep
  output_varIMPtotal.pm <- rbind(output_varIMPtotal.pm, varIMP.pm)

if( (rep %% 10) == 0){ # take rep, divide by 10, outputs remainder
  print(paste("Done with", rep, "step"))
} 

}

#7.
#Overall correct classification scores
correctP.pm <- round((results_pm[1,1]/(results_pm[1,1]+results_pm[1,2])*100), 2)
correctM.pm <- round((results_pm[2,2]/(results_pm[2,1]+results_pm[2,2])*100), 2)

output_CorrectClass.pm <- rbind(results_pm,cbind(correctP.pm, correctM.pm))

names(pc_classPM) <- c("rep", "EncounterID", "population", "classification")



proc.time() - ptm

```

####Calculate means and standard errors
```{r}
se <- function(x) sd(x)/sqrt(length(x))

meanRealP.pm <- mean(realP.pm)
meanRealM.pm <- mean(realM.pm)
seRealP.pm <- se(realP.pm)
seRealM.pm <- se(realM.pm)

 
date='20190528'
#Combine means and std errors into matrix
sink("data\\results\\ConfMatrix_PM_means_20190528.txt", append=T)
PNMmeans.pm <- cbind("PEL"= c(meanRealP.pm), "MHI"=c(meanRealM.pm)) #means
PNMse.pm <- cbind("PEL"= c(seRealP.pm), "MHI"=c(seRealM.pm)) #standard errors
PNMtotal.pm <- rbind("Means" = PNMmeans.pm, "Stand Err" = PNMse.pm)

date
w
n_samps
PNMtotal.pm
sink()


```

#save to project folder
```{r}

#Save overall confusion matrix
write.table(output_CorrectClass.pm, "data\\results/PcResults_PM_ConfusionMatrix_20190528.csv", append = T, row.names = T, col.names=T, sep = ",")

write.table(output_results_pm, "data\\results\\PcResults_PM_TotalResults_20190528.csv", append = T, row.names = T, col.names=T, sep = ",")

#write.table(output_WhistleClass.pm, "data\\results\\PcResults_PM_TotalWhistleClass_20190325.csv", append = T, row.names = F, col.names=T, sep = ",")

write.table(output_TrainSetAll.pm, "data\\results\\PcResults_PM_TrainSets_20190528.csv", append = T, col.names=T, sep = ",")

write.table(output_TestSetAll.pm, "data\\results\\PcResults_PM_TestResults_20190528.csv", append = T, col.names=T, sep = ",")

write.table(output_varIMPtotal.pm, "data\\results\\PcResults_PM_ImportantVars_20190528.csv", append = T, col.names=T, sep = ",")

write.table(tunedAll.pm, "data\\results\\PcResults_PM_TunedParameters_20190528.csv", append = T, col.names=T, sep = ",")

write.table(output_ModelFit.pm, "data\\results\\PcResults_PM_TotalModels_20190528.csv", append = T, col.names=F, sep = ",")


write.table(pc_classPM, "data/results/PcResults_PM_EncounterClass_20190528.csv", append = F, col.names=T, row.names = F, sep = ",")

```

#For laptop directories

```{r}
se <- function(x) sd(x)/sqrt(length(x))

meanRealP <- mean(realP)
meanRealM <- mean(realM)
seRealP <- se(realP)
seRealM <- se(realM)

 
date='20180822'
#Combine means and std errors into matrix
sink("C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018\\ConfMatrixPM_means_20180822.txt", append=T)
PNMmeans.pm <- cbind("PEL"= c(meanRealP), "MHI"=c(meanRealM)) #means
PNMse.pm <- cbind("PEL"= c(seRealP), "MHI"=c(seRealM)) #standard errors
PNMtotal.pm <- rbind("Means" = PNMmeans.pm, "Stand Err" = PNMse.pm)

date
w
n_samps
PNMtotal.pm
sink()


```


###Write NEW WHISTLE DATA outputs to file
```{r}

#Save overall confusion matrix
write.table(output_CorrectClass.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//20180818//PM//PcResults_ConfusionMatrix_PM_20180829_noMHI34.csv", append = T, row.names = T, col.names=T, sep = ",")

write.table(output_results_pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//20180818//PM//PcResults_TotalResults_PM_20180818.csv", append = T, row.names = T, col.names=T, sep = ",")

write.table(output_WhistleClass.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//20180818//PM//PcResults_TotalWhistleClass_PM_20180818.csv", append = T, row.names = F, col.names=T, sep = ",")

write.table(output_TrainSetAll.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//20180818//PM///Top5ImpVar_PcResults_TrainSets_newPvM_NoPel8.csv", append = T, col.names=T, sep = ",")

write.table(output_TestSetAll.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//20180818//PM//Top5ImpVar_PcResults_TestResults_newPvM_NoPel8.csv", append = T, col.names=T, sep = ",")

write.table(output_varIMPtotal.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//20180818//PM//Top5ImpVar_PcResults_ImportantVars_newPvM_NoPel8.csv", append = T, col.names=T, sep = ",")

write.table(tunedAll.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//20180818//PM//Top5ImpVar_PcResults_TunedParameters_newPvM_NoPel8.csv", append = T, col.names=T, sep = ",")

#9) Save encounter classifications based on majority of whistles
write.table(pc_classPM, "data/results/PcResults_PM_EncounterClass_20190322.csv", append = F, col.names=T, row.names = F, sep = ",")

```


#Variance 
```{r}
#Load file with all confusion matrices for each trial
my_data <- read.csv('data/results/PcResults_TotalResults_PM_20190325.csv')

#Create separate objects with correct scores for each population
my_data_p <- filter(my_data, X == 'Pelagic' ) 
my_data_p <- my_data_p[,-1]
my_data_m <- filter(my_data, X == 'MHI')
my_data_m <- my_data_m[,-1]

#Variance of correct classifications (diagonal)
my_var_p <- var(apply(my_data_p, 1, function(x) x[1]/sum(x[1],x[2])))
my_var_m <- var(apply(my_data_m, 1, function(x) x[2]/sum(x[1],x[2])))

#Variance of Pelagic misclassifications 
my_var_pm <- var(apply(my_data_p, 1, function(x) x[2]/sum(x[1],x[2])))

#Variance of MHI misclassifications
my_var_mp <- var(apply(my_data_m, 1, function(x) x[1]/sum(x[1],x[2])))

```