This code plots the distribution of the ROCCA variables for each population to explore the data and know what we're working with for statistical tests.


```{r message=FALSE}
library(dplyr)

pcplot<- read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/ROCCA Towed Data/BigRoccaFile_PMN_TowedAll_EDIT_20180914.csv')
pcdata.P <- droplevels(filter(pcplot, population == "Pelagic"))      
pcdata.N <- droplevels(filter(pcplot, population == "NWHI"))      
pcdata.M <- droplevels(filter(pcplot, population == "MHI"))

#remove columns
pcplotsub <- pcplot[, -c(2:6, 54:61)]
pcdata.Msub <- pcdata.M[, -c(2,4:6, 54:61)]
names<-names(pcdata.Msub)
names<-names(pcplotsub)
names<-names[-c(1)]
```

##Density plots of variables across populations
```{r}
### 9/12/2018 ####
library(ggplot2)
library(reshape2)
library(ggpubr)

pcplotsub$population <-factor(pcplot$population, levels=c("Pelagic", 'NWHI', 'MHI'), ordered = T) #reorder population factor
names<-names(pcplotsub)
names<-names[-1]
# #This works below, but plots separate histograms if the data are subsetted for each population beforehand
# classes<-sapply(pcdata.Msub,class)
# names[classes == 'numeric']
# par(mfrow=c(4, 4))
# for (name in names[classes == 'numeric']){
#  
#  p <- hist(pcdata.Msub[,name], main = name)
#   ggsave 
# }


##Functioning code for Density Plots in multipanels on multiple pages ####
  #ggplot2 density plots
plots <- list()
for(nm in names){
plots[[nm]]<- ggplot(pcplotsub, aes_string(x=nm, fill=pcplotsub$population)) +  geom_density(alpha = 0.3) +
    theme_bw()+
    theme(plot.background = element_rect(fill = "white"),
        axis.title.x=element_text(size=10, face='bold'),
        axis.text.x=element_text(size=8),
        axis.text.y=element_text(size=8)) +
        guides(fill=guide_legend(title="Population"))
#ggsave(filename = paste("PcDensityPlot", nm, '.png', sep = ""), width = 12, height = 8, units = "in", type = "cairo-png")
}
#arrange plots 
multipage <- ggarrange(plotlist=plots, ncol = 2, nrow = 4, common.legend = TRUE, legend = "bottom" )
#export plots to multiple pages
ggexport(multipage, filename = "multi.page.ggplot2.pdf")
#another option to plot on multiple pages, but not as good
plots2 = dlply(pcplot , names, `%+%`, e1 = plots[[nm]])
ml = marrangeGrob(plots, nrow=4, ncol=2)
ggsave("multipage.pdf", ml)


### By encounter
#change factor order
#reverse
#  pcdata.Msub$population = with(pcdata.Msub, factor(population, levels = rev(levels(population))))
#custom
levels(pcdata.Msub$EncounterID) <- c("M1", "M2", "M3", "M4" )

plots.P <- list()
plots.N <- list()
plots.M <- list()
for(nm in names){
plots.P[[nm]]<- ggplot(pcdata.P, aes_string(x=nm, fill=pcdata.P$EncounterID)) +  geom_density(alpha = 0.3) +
    theme_bw()+
    theme(plot.background = element_rect(fill = "white"),
        axis.title.x=element_text(size=10, face='bold'),
        axis.text.x=element_text(size=8),
        axis.text.y=element_text(size=8)) +
        guides(fill=guide_legend(title="Population"))
#ggsave(filename = paste("PcDensityPlot", nm, '.png', sep = ""), width = 12, height = 8, units = "in", type = "cairo-png")
}
#arrange plots 
multipage <- ggarrange(plotlist=plots.P, ncol = 2, nrow = 4, common.legend = TRUE, legend = "bottom" )
#export plots to multiple pages
ggexport(multipage, filename = "PcDensityPlot_PelagicVariables.pdf")


```


## Mean of the Mean Positive Slope for MHI
```{r}
df_mhi = droplevels(subset(pcplot, pcplot$population=='mhi'))
posslope_mhi = mean(df_mhi$FREQPOSSLOPEMEAN)
plot(as.factor(df_mhi$group), df_mhi$FREQPOSSLOPEMEAN, xlab="Encounter ID", ylab="Mean Positive Slope", main="Main Hawaiian Island Population")
plot(as.factor(df_mhi$group), df_mhi$DURATION, xlab="Encounter ID", ylab="Mean Positive Slope", main="Main Hawaiian Island Population")
plot(as.factor(df_mhi$group), df_mhi$FREQSWEEPFLATPERCENT, xlab="Encounter ID", ylab="% of the whistle with zero slope ", main="Main Hawaiian Island Population")

posslope_mhi = mean(df_mhi$FREQNEGSLOPEMEAN)
plot(df_mhi$id, df_mhi$FREQNEGSLOPEMEAN, xlab="Encounter ID", ylab="Mean Negative Slope", main="Main Hawaiian Island Population")

plot(df_mhi$id, df_mhi$DURATION, xlab="Encounter ID", ylab="Mean Negative Slope", main="Main Hawaiian Island Population")
plot(df_mhi$id, df_mhi$FREQSWEEPFLATPERCENT, xlab="Encounter ID", ylab="Mean Negative Slope", main="Main Hawaiian Island Population")
plot(df_mhi$id, df_mhi$FREQSLOPEMEAN, xlab="Encounter ID", ylab="Mean Negative Slope", main="Main Hawaiian Island Population")


mhi30 = filter(df_mhi, df_mhi$group == 30)
#Mean of mean positive slope for each  mhi encounter 
write.csv(list(aggregate(df_mhi$FREQPOSSLOPEMEAN, by=list(df_mhi$id), FUN=mean)),'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\MHI_MeanPosSlope.csv') 
aggregate(df_mhi$FREQNEGSLOPEMEAN, by=list(df_mhi$id), FUN=mean)
aggregate(df_mhi$FREQPOSSLOPEMEAN, by=list(df_mhi$Recorder), FUN=median)
aggregate(df_mhi$DURATION, by=list(df_mhi$Recorder), FUN=median)
```

mean(mhi30redo$FREQPOSSLOPEMEAN)

library(lattice)
## Mean of the Mean Positive Slope for PELAGIC
```{r}
df_pel = droplevels(subset(pcplot, pcplot$population=='pelagic'))
posslope_pel = mean(df_pel$FREQPOSSLOPEMEAN)
plot(as.factor(df_pel$group), df_pel$FREQPOSSLOPEMEAN, xlab="Encounter ID", ylab="Mean Positive Slope", main= "Pelagic Population")

posslope_pel = mean(df_pel$FREQNEGSLOPEMEAN)
plot(as.factor(df_pel$group), df_pel$FREQNEGSLOPEMEAN, xlab="Encounter ID", ylab="Mean Negative Slope", main= "Pelagic Population")

plot(as.factor(df_pel$group), df_pel$FREQSWEEPFLATPERCENT, xlab="Encounter ID", ylab="% of the whistle with zero slope ", main="Pelagic Population")
plot(as.factor(df_pel$group), df_pel$FREQQUARTER3, xlab="Encounter ID", ylab="% of the whistle with zero slope ", main="Pelagic Population")

write.csv(list(aggregate(df_pel$FREQPOSSLOPEMEAN, by=list(df_pel$id), FUN=mean)),'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\Pelagic_MeanPosSlope.csv') 
aggregate(df_pel$FREQNEGSLOPEMEAN, by=list(df_pel$id), FUN=mean)
aggregate(df_pel$FREQPOSSLOPEMEAN, by=list(df_pel$Recorder), FUN=median)
aggregate(df_pel$FREQQUARTER3, by=list(df_pel$id), FUN=mean)
```

## Mean of the Mean Positive Slope for NWHI
```{r}
df_nwhi = droplevels(subset(pcplot, pcplot$population=='nwhi'))
posslope_nwhi = mean(df_nwhi$FREQPOSSLOPEMEAN)
plot(as.factor(df_nwhi$group), df_nwhi$FREQPOSSLOPEMEAN, xlab="Encounter ID", ylab="Mean Positive Slope", main="Northwestern Hawaiian Island Population")

posslope_nwhi = mean(df_nwhi$FREQNEGSLOPEMEAN)
plot(df_nwhi$id, df_nwhi$FREQNEGSLOPEMEAN, xlab="Encounter ID", ylab="Mean Negative Slope", main="Northwestern Hawaiian Island Population")


plot(as.factor(df_nwhi$group), df_nwhi$FREQSWEEPFLATPERCENT, xlab="Encounter ID", ylab="% of the whistle with zero slope ", main="Northwestern Hawaiian Island Population")

write.csv(list(aggregate(df_nwhi$FREQPOSSLOPEMEAN, by=list(df_nwhi$id), FUN=mean)),'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\NWHI_MeanPosSlope.csv')
aggregate(df_nwhi$FREQPOSSLOPEMEAN, by=list(df_nwhi$id), FUN=mean)
aggregate(df_nwhi$FREQNEGSLOPEMEAN, by=list(df_nwhi$id), FUN=mean)
aggregate(df_nwhi$FREQPOSSLOPEMEAN, by=list(df_nwhi$Recorder), FUN=median)
aggregate(df_nwhi$FREQQUARTER3, by=list(df_nwhi$id), FUN=mean)


plot(as.factor(pcplot$group), pcplot$FREQQUARTER3, col=pcplot$population, xlab="Encounter ID", ylab="Frequency of the third quarter of the whistle")
```


```{r}
library(ggplot2)
#shades of gray###
color <- c('#35bb7c', 'black', '#947d51','#8576aa')


#Plot Mean positive slope / duration
freqslope=ggplot(pcplot, aes(x = EncounterID, y = FREQPOSSLOPEMEAN, col=as.factor(population))) +
  geom_boxplot() +
  theme_bw() +
  labs(x="Encounter", y="Frequency (Hz)", 
       title="Mean Positive Slope") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=7),
        axis.text.y = element_text(size=9),
        axis.title=element_text(size=12,face="bold"),
        legend.title = element_blank()) +
  scale_color_manual(values=color) 
ggsave('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\Variable Importance\\Plot_FREQPOSSLOPE.png', width = 11, height = 8, units = "in", type = "cairo-png")




# Plot third-quarter frequency / duration
freqspd<-ggplot(pcplot, aes(x = EncounterID, y = FREQSPREAD, col=as.factor(population))) +
  geom_boxplot() +
  theme_bw() +
  labs(x="Encounter", y="Frequency (Hz)")+ 
       #title="Third Quarter Frequency", 
       #subtitle="Frequency at three-quarters of the duration") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=7),
        axis.text.y = element_text(size=9),
        axis.title=element_text(size=12,face="bold"),
        legend.title = element_blank()) +
  scale_color_manual(values=color) 
 
  

ggsave('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\Variable Importance\\Plot_freqspread.png', width = 11, height = 8, units = "in", type = "cairo-png")




# Plot FREQSWEEPFLATPERCENT
freq <- ggplot(pcplot, aes(x = EncounterID, y = pcplot$FREQQUARTER, col=as.factor(population))) +
  geom_boxplot() +
  theme_bw() +
  labs(x="Encounter", y="", 
       title="Percent Flat",
       subtitle="Percent of the whistle with zero slope") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=7),
        axis.text.y = element_text(size=9),
        axis.title=element_text(size=12,face="bold"),
        legend.title = element_blank()) +
  guides(shape = guide_legend(override.aes = list(size=10)))+
  scale_color_manual(values=color)
ggsave('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\Variable Importance\\Plot_FREQQUARTER.png', width = 11, height = 8, units = "in", type = "cairo-png")
#ggsave('Plot_FREQSWEEPFLAT.png', device='pdf', width = 11, height = 8, units = "in")
```

```{r}
# Plot Accuracy by Duration
pcACC <- read.csv("C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data/Barkley_Sensitivity Analysis Table.csv")

#purple, green, orange
colorPOP <- c('#791fb8','#2b961a','darkorange2')

ggplot(pcACC, aes(x = pcACC$duration..s., y = accuracy, col=as.factor(Recorder))) +
  
  theme_bw() +
  labs(x="Duration (s)", y="Percent Accuracy")+ 
  geom_text(aes(label=pcACC$EncounterPoster), size=5, vjust=-0.1)+
  geom_point() +
  #title="Percent Flat", 
  #subtitle="Percent of the whistle with zero slope") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12),
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_blank(),
        legend.text=element_text(size=15)) +
  guides(shape=guide_legend(override.aes = list(shape = 16)))+
  scale_color_manual(values=color) 
ggsave('Plot_AccVSDur.pdf', width = 11, height = 8, units = "in")


#just to get legend with dots

ggplot(pcACC, aes(x = pcACC$duration..s., y = accuracy, col=as.factor(Recorder))) +
  
  theme_bw() +
  labs(x="Duration (s)", y="Percent Accuracy")+ 
  geom_point() +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12),
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_blank(),
        legend.text=element_text(size=15)) +
  guides(color = guide_legend(override.aes = list(size=5)))+
  scale_color_manual(values=color)
ggsave('Plot_AccVSDurLEGEND.pdf', width = 11, height = 8, units = "in")


ggplot(pcACC, aes(x = pcACC$Recorder, y = accuracy, col=as.factor(Population))) +
  
  theme_bw() +
  labs(x="Recorder Type", y="Percent Accuracy")+ 
  geom_point()+
  geom_text(aes(label=pcACC$EncounterPoster), size=4, hjust=0, nudge_x=0.05)+
  #title="Percent Flat", 
  #subtitle="Percent of the whistle with zero slope") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        plot.title = element_text(size=20, hjust=0.5, face='bold'),
        plot.subtitle=element_text(size=15, hjust=0.5, face="italic", color="black"),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size=12, face='bold'),
        axis.text.y = element_text(size=12, face='bold'),
        axis.title=element_text(size=16,face="bold"),
        legend.position = "none") + #no legend
  guides(shape=guide_legend(override.aes = list(shape = 16)))+
  scale_color_manual(values=colorPOP) 
ggsave('Plot_AccVSDur.png', width = 11, height = 8, units = "in", type = "cairo-png")
```

```{r}
#Explore Cluster 4, MHI 29

mhi29 <- subset(pcplot, pcplot$group == 30)

mhi.only <- subset(pcplot, pcplot$population == 'MHI')
```








```{r}

for (i in colnames) {
    hist(pcdata.M[,i], xlim=c(0, 12000), breaks=seq(0, 12000, 100), main=colnames[i], probability=TRUE, col="gray", border="white")
}

#violin plot
library(vioplot)

colnames <-colnames(pcplot)[4:50]

par(mfrow=c(3,3))

  ggplot(pcplot, aes(x=population, y=FREQMAX)) + geom_violin()
  ggplot(pcplot, aes(x=population, y=FREQMIN)) + geom_violin()


melted = melt(pcplot, id =c("population")) 

ggplot(melted, aes(value)) + facet_grid(~variable) + geom_histogram()+
  theme(axis.text.x=element_text(size=8), 
        axis.title.x=element_text(size=8, face='bold'),
        axis.text.y=element_text(size=8), 
        axis.title.y=element_text(size=8, face='bold'),
         text=element_text(size=8),
        plot.title=element_text(size=12, hjust = 0.5, face = 'bold'),
        strip.text.x = element_text(size = 5),
        plot.margin = unit(c(0.5,0.25,0.5,0.25), "cm"))





```