---
title: "Kolmogorov-Smirnov Tests"
author: "Yvonne Barkley"
date: "November 25, 2018"
output: html_document
---


```{r}
library(dplyr)
library(data.table)

pcdata <- read.csv('C:\\Users\\yvers\\Documents\\CHP 1\\data\\ROCCA Towed Data\\BigRoccaFile_PMN_TowedAll_EDIT_20181124.csv')

## Pull out continuous variables
pcdata_cont <- pcdata[, c(1:24,28:32,45:47,49:53)]

#Filter out correlated variables
library(corrplot)
pcsub<-pcdata_cont[,7:37]
corrplot.mixed(cor(pcsub, method= "pearson"), upper="number", lower="circle")
tmp <- cor(pcsub)
tmp[upper.tri(tmp)] <- 0
diag(tmp) <- 0
pcsubcor <- pcsub[,!apply(tmp,2,function(x) any(x > 0.80 | x < -0.80))]
pcdata_cont2 <- cbind(pcdata_cont[, c(1:6)], pcsubcor)


# pcdata.MN <- filter(pcdata, population == "MHI" | population == "NWHI")
# pcdata.PM <- droplevels(filter(pcdata, population == "MHI" | population == "Pelagic"))
# pcdata.PN <- droplevels(filter(pcdata, population == "Pelagic" | population == "NWHI"))


pcdata.P <- droplevels(filter(pcdata_cont, population == "Pelagic"))      
pcdata.N <- droplevels(filter(pcdata_cont, population == "NWHI"))      
pcdata.M <- droplevels(filter(pcdata_cont, population == "MHI"))

```





## KS test - Oct 5, 2018 Annie's suggestion ##


BETWEEN POPULATIONS
Test correlation of variables for each POPULATION separately, then match up the variables that are left.

##MHI correlation
```{r}    
#Using only continuous variables, determine the correlated variables within the MHI whistles
pcsub<-pcdata.M[,7:37]
corrplot.mixed(cor(pcsub, method= "pearson"), upper="number", lower="circle")
tmp <- cor(pcsub)
tmp[upper.tri(tmp)] <- 0
diag(tmp) <- 0
pcsubcor_M <- pcsub[,!apply(tmp,2,function(x) any(x > 0.80 | x < -0.80))]

pcdata.Muncorr <- cbind(pcdata.M[, c(1:6)], pcsubcor_M)

```

##NWHI correlation
```{r}    
#Using only continuous variables, determine the correlated variables within the NWHI whistles
pcsub<-pcdata.N[,7:37]
#corrplot.mixed(cor(pcsub, method= "pearson"), upper="number", lower="circle")
tmp <- cor(pcsub)
tmp[upper.tri(tmp)] <- 0
diag(tmp) <- 0
pcsubcor_N <- pcsub[,!apply(tmp,2,function(x) any(x > 0.80 | x < -0.80))]

pcdata.Nuncorr <- cbind(pcdata.N[, c(1:6)], pcsubcor_N)

```

##Pelagic correlation
```{r}    
##Using only continuous variables, determine the correlated variables within the Pelagic whistles
pcsub<-pcdata.P[,7:37]
#corrplot.mixed(cor(pcsub, method= "pearson"), upper="number", lower="circle")
tmp <- cor(pcsub)
tmp[upper.tri(tmp)] <- 0
diag(tmp) <- 0
pcsubcor_P <- pcsub[,!apply(tmp,2,function(x) any(x > 0.80 | x < -0.80))]

pcdata.Puncorr <- cbind(pcdata.P[, c(1:6)], pcsubcor_P)

```


##Run KS Tests
```{r}
#create dataframe combining uncorrelated variables and population column
pcsubcor_M <- cbind('population' = pcdata.Muncorr[,1], pcdata.Muncorr[,c(7:length(pcdata.Muncorr))])
pcsubcor_N <- cbind('population' = pcdata.Nuncorr[,1], pcdata.Nuncorr[,c(7:length(pcdata.Nuncorr))])
pcsubcor_P <- cbind('population' = pcdata.Puncorr[,1], pcdata.Puncorr[,c(7:length(pcdata.Puncorr))])




#find the common uncorrelated variables matching between each pair of populations
vars_M = names(pcsubcor_M)
vars_N = names(pcsubcor_N)
vars_P = names(pcsubcor_P)

##MHI vs NWHI##
# These are the variables in NWHI that match with MHI - omit na's or the vector won't be the proper length
idx_M <- na.omit(match(vars_M, vars_N))
# idx_M <- match(names(pcdata.Muncorr[ ,c(7:length(pcdata.Muncorr))]), names(pcdata.Nuncorr[, c(7:length(pcdata.Nuncorr))]), nomatch = NA_integer_, incomparables = NULL )

#These are the variables in MHI that match with MHI - omit na's or the vector won't be the proper length
idx_N <- na.omit(match(vars_N, vars_M))

# idx_N <- na.omit(match(names(pcdata.Nuncorr[ ,c(7:length(pcdata.Nuncorr))]), names(pcdata.Muncorr[, c(7:length(pcdata.Muncorr))])) )

ks_vars_m <- pcsubcor_M[, idx_N]
ks_vars_n <- pcsubcor_N[, idx_M]

#Remove population column before KS test
ks_vars_m <- ks_vars_m[, -1]
ks_vars_n <- ks_vars_n[, -1]

####MHI vs NWHI####
ks_resultsMN = NULL
#names_ks <-names(ks_vars_m)

for (i in colnames(ks_vars_m))
     {
  #for (n in colnames(ks_vars_n))
     #{
       ksMN <- ks.test(ks_vars_m[,i], ks_vars_n[,i])
       
      ksMN_df <- cbind(colnames(ks_vars_m[i]),ksMN$statistic,ksMN$p.value)

      ks_resultsMN <- rbind(ks_resultsMN, ksMN_df)
    
       }
  colnames(ks_resultsMN) <- c("variable", "statistic", "p.value")   
comparison = 'NWHI-MHI'
ks_resultsMN <- cbind(ks_resultsMN, comparison) 


####MHI vs Pelagic####
idx_M2 <- na.omit(match(vars_M, vars_P)) 
idx_P <- na.omit(match(vars_P, vars_M))

ks_vars_m2 <- pcsubcor_M[, idx_P] 
ks_vars_p <- pcsubcor_P[, idx_M2]
 
#Remove population column before KS test
ks_vars_m2 <- ks_vars_m2[, -1]
ks_vars_p <- ks_vars_p[, -1]
  
  
ks_resultsMP = NULL
#names_ks <-names(ks_vars_m)

for (i in colnames(ks_vars_m2))
     {
  #for (n in colnames(ks_vars_n))
     #{
       ksMP <- ks.test(ks_vars_m2[,i], ks_vars_p[,i])
       
      ksMP_df <- cbind(colnames(ks_vars_m2[i]), ksMP$statistic, ksMP$p.value)

      ks_resultsMP <- rbind(ks_resultsMP, ksMP_df)
    
       }
  colnames(ks_resultsMP) <- c("variable", "statistic", "p.value") 
comparison = 'Pelagic-MHI'
ks_resultsMP <- cbind(ks_resultsMP, comparison)  

####Pelagic vs NWHI####
idx_P <- na.omit(match(vars_P, vars_N))
idx_N2 <- na.omit(match(vars_N, vars_P)) 

ks_vars_p <- pcsubcor_P[, idx_N2]
ks_vars_n2 <- pcsubcor_N[, idx_P] 

#Remove population column before KS test
ks_vars_p <- ks_vars_p[, -1]
ks_vars_n2 <- ks_vars_n2[, -1]

ks_resultsPN = NULL
#names_ks <-names(ks_vars_m)

for (i in colnames(ks_vars_p))
     {
  #for (n in colnames(ks_vars_n))
     #{
       ksPN <- ks.test(ks_vars_p[,i], ks_vars_n2[,i])
       
      ksPN_df <- cbind(colnames(ks_vars_p[i]), ksPN$statistic, ksPN$p.value)

      ks_resultsPN <- rbind(ks_resultsPN, ksPN_df)
    
       }
  colnames(ks_resultsPN) <- c("variable", "statistic", "p.value") 
  
  comparison = 'Pelagic-NWHI'
  ks_resultsPN <- cbind(ks_resultsPN, comparison)
```


##Format KS dataframe  
```{r}  
#Combine all ks results into one data frame, pvalues are factors!
ks <- as.data.frame(rbind(ks_resultsMN, ks_resultsMP, ks_resultsPN))
ks$p.value <- as.numeric(levels(ks$p.value))[ks$p.value]

ks$variable <- as.character(ks$variable)
#levels(ks$comparison) <- c('NWHI-MHI','Pelagic-MHI','Pelagic-NWHI')

#change comparison labels and order the levels
# ks$comparison[grep("MN", ks$comparison)] <- "NWHI-MHI"
# ks$comparison[grep("PN", ks$comparison)] <- "Pelagic-NWHI"
# ks$comparison[grep("MP", ks$comparison)] <- "Pelagic-MHI"
# 
# ks$comparison <- factor(as.character(ks$comparison), levels=c("Pelagic-NWHI", 'Pelagic-MHI', 'NWHI-MHI'))

##For plotting all variables to test significance
ks$variable[grep("DURATION", ks$variable)] <- "Duration"
ks$variable[grep("FREQQUARTER3", ks$variable)] <- "ThirdQuarterFreq"
ks$variable[grep("FREQSPREAD", ks$variable)] <- "FreqSpread"
ks$variable[grep("FREQCOFM", ks$variable)] <- "CoeffFreqMod"
#ks$variable[grep("FREQMAXMINRATIO", ks$variable)] <- "MaxMinRatio"
ks$variable[grep("FREQSLOPEMEAN", ks$variable)] <- "MeanSlope"
ks$variable[grep("FREQABSSLOPEMEAN", ks$variable)] <- "AbsSlopeMean"
ks$variable[grep("FREQPOSSLOPEMEAN", ks$variable)] <- "MeanPosSlope"
ks$variable[grep("FREQNEGSLOPEMEAN", ks$variable)] <- "MeanNegSlope"
ks$variable[grep("FREQSLOPERATIO", ks$variable)] <- "PosNegSlopeRatio"
# ks$variable[grep("FREQBEGDWN", ks$variable)] <- "BegDwnFreq"
# ks$variable[grep("FREQENDUP", ks$variable)] <- "UpEndFreq"
# ks$variable[grep("FREQENDDWN", ks$variable)] <- "EndDwnFreq"
# ks$variable[grep("NUMSWEEPSFLATUP", ks$variable)] <- "NumTimesZeroToPos"
ks$variable[grep("FREQSWEEPDWNPERCENT", ks$variable)] <- "PctNegSlope"
ks$variable[grep("FREQSWEEPFLATPERCENT", ks$variable)] <- "PctZeroSlope"
#ks$variable[grep("NUMSWEEPSFLATDWN", ks$variable)] <- "NumTimesZeroToNeg"
#ks$variable[grep("NUMINFLECTIONS", ks$variable)] <- "NumInfl"
#ks$variable[grep("FREQBEGENDRATIO", ks$variable)] <- "BegEndRatioFreq"
ks$variable[grep("INFLMAXMINDELTA", ks$variable)] <- "MaxMinInflTimeRatio"
ks$variable[grep("INFLMEANDELTA", ks$variable)] <- "MeanTimeBtwnInfl"
ks$variable[grep("INFLSTDDEVDELTA", ks$variable)] <- "SDTimeBtwnInfl"




write.csv(ks, 'data/results/KStests_ALL_20190318.csv', row.names = F)
```

```{r}
#filter for p-values < 0.05
ks_sig <- filter(ks, ks$p.value < 0.05)
ks_sig$variable <-as.character(ks_sig$variable)


ks_sig$variable[grep("DURATION", ks_sig$variable)] <- "Duration"
ks_sig$variable[grep("FREQQUARTER3", ks_sig$variable)] <- "ThirdQuarterFreq"
ks_sig$variable[grep("FREQSPREAD", ks_sig$variable)] <- "FreqSpread"
ks_sig$variable[grep("FREQCOFM", ks_sig$variable)] <- "CoeffFreqMod"
ks_sig$variable[grep("FREQPOSSLOPEMEAN", ks_sig$variable)] <- "MeanPosSlope"
ks_sig$variable[grep("FREQNEGSLOPEMEAN", ks_sig$variable)] <- "MeanNegSlope"
ks_sig$variable[grep("FREQSWEEPDWNPERCENT", ks_sig$variable)] <- "PctNegSlope"
ks_sig$variable[grep("INFLMAXMINDELTA", ks_sig$variable)] <- "MaxInflTime/MinInflTime"
ks_sig$variable[grep("INFLMEANDELTA", ks_sig$variable)] <- "MeanTimeBtwnInfl"
ks_sig$variable[grep("INFLSTDDEVDELTA", ks_sig$variable)] <- "SDTimeBtwnInfl"
ks_sig$variable[grep("FREQSWEEPFLATPERCENT", ks_sig$variable)] <- "PctZeroSlope"
ks_sig$variable[grep("INFLMEANDELTA", ks_sig$variable)] <- "MeanTimeBtwnInfl"
ks_sig$variable[grep("FREQABSSLOPEMEAN", ks_sig$variable)] <- "AbsSlopeMean"
ks_sig$variable[grep("NUMSWEEPSFLATUP", ks_sig$variable)] <- "NumTimesZeroToPos"
ks_sig$variable[grep("FREQSLOPEMEAN", ks_sig$variable)] <- "MeanSlope"
ks_sig$variable[grep("NUMSWEEPSFLATUP", ks_sig$variable)] <- "NumTimesZeroToPos"
ks_sig$variable[grep("FREQENDDWN", ks_sig$variable)] <- "BegDwnFreq"
ks_sig$variable[grep("FREQMAXMINRATIO", ks_sig$variable)] <- "MaxMinRatio"
ks_sig$variable[grep("NUMSWEEPSFLATDWN", ks_sig$variable)] <- "NumTimesZeroToNeg"
ks_sig$variable[grep("NUMINFLECTIONS", ks_sig$variable)] <- "NumInfl"
ks_sig$variable[grep("FREQMAXMINRATIO", ks_sig$variable)] <- "MaxMinRatio"
ks_sig$variable[grep("FREQBEGDWN", ks_sig$variable)] <- "BegDwnFreq"
ks_sig$variable[grep("FREQBEGENDRATIO", ks_sig$variable)] <- "BegEndRatioFreq"
ks_sig$variable[grep("FREQENDUP", ks_sig$variable)] <- "UpEndFreq"
ks_sig$variable[grep("FREQSLOPERATIO", ks_sig$variable)] <- "PosNegSlopeRatio"


write.csv(ks_sig, 'C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results/2018/20181125_KS Tests/KStests_SIGNIF_ALL.csv', row.names = F)
```


#Plot KS test results
```{r}
## Make plot showing how the variables differ between populations

# Load data
 ks <- read.csv('data/results/KStests_ALL_20190318.csv')
library(ggplot2)

#Since taking the log of the pvalues, log(0) = inf, which creates problems for plot.
#Here, replace zero with 1e-22
ks$p.value <-as.character(ks$p.value)
ks$p.value[ks$p.value == 0] <- '1e-22'
ks$p.value <-as.numeric(ks$p.value)
ks$comparison <- factor(ks$comparison, levels = c("Pelagic-NWHI","Pelagic-MHI" ,"NWHI-MHI"))


# require(ggplot2)

p_ks <- ggplot(ks, aes(p.value, variable)) + 
     geom_point(aes(pch=factor(comparison)), size=4.5)+
     
  labs(title="Kolmogorov-Smirnov Test Results", x = "p-value", y= 'Whistle Variable') + 
  xlim(-1, 1)+
  theme_bw()+
  theme(plot.background = element_rect(fill = "white"),
        axis.text.x=element_text(size=12, angle = 90, hjust = 1), #5 for all encounters
        axis.text.y=element_text(size=12), #8 for all encounters
        plot.title=element_text(size=16, hjust=0.5, face="bold", colour="black", vjust=-1),
        legend.text = element_text(size=12),
        legend.title = element_text(size=14),
        legend.title.align = 0.5,
        axis.title = element_text(size=13, hjust=0.5))+
   #use inkscape to mask stupid size legend
    scale_shape_manual(values = c(20,1,8),
                     name='Comparison')+ 
  scale_x_log10()#(breaks = c(0,1e-10,1e-5,.05,.75,1)) 
#scale_color_discrete(name='Model') #only works for colored points
p_ks + guides(shape = guide_legend(override.aes = list(size = 5)))+
       geom_vline(xintercept=0.05, linetype='dashed', color='red', size=1)

ggsave('data/results/KS_plot_20190424.png', width = 12, height = 8, units = "in", type = "cairo-png")

#ggsave('C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018\\plots/KS TEST_20190319-new.png', width = 12, height = 8, units = "in", type = "cairo-png")  

#plot(ks$p.value, ks$variable, log='x' )


```


```{r}
# Make two random samples
sample1<-rnorm(1000)
sample2<-rnorm(1000, 2, 2)

group <- c(rep("sample1", length(sample1)), rep("sample2", length(sample2)))
dat <- data.frame(KSD = c(sample1,sample2), group = group)
cdf1 <- ecdf(sample1) 
cdf2 <- ecdf(sample2) 

minMax <- seq(min(sample1, sample2), max(sample1, sample2), length.out=length(sample1)) 
x0 <- minMax[which( abs(cdf1(minMax) - cdf2(minMax)) == max(abs(cdf1(minMax) - cdf2(minMax))) )] 
y0 <- cdf1(x0) 
y1 <- cdf2(x0) 


ggplot(dat, aes(x = KSD, group = group, colour = group, linetype=group))+
  stat_ecdf(size=1) +
  xlab("mm") +
  ylab("Cumulitive Distibution") +
  geom_segment(aes(x = x0[1], y = y0[1], xend = x0[1], yend = y1[1]),
               linetype = "dashed", color = "red") +
  geom_point(aes(x = x0[1] , y= y0[1]), color="red", size=1) +
  geom_point(aes(x = x0[1] , y= y1[1]), color="red", size=1) +
  ggtitle("K-S Test: Sample 1 / Sample 2")
```





Test for normality in uncorrelated variables using Shapiro-Wilks test with Bonferroni Correction and with Normal Q-Q plots
###only duration and the third quarter frequency were normal
```{r}
qqnorm(ks_vars_m$DURATION)

shapiro_test_df <- function(df, bonf= TRUE, alpha= 0.05) {
  l <- lapply(df, shapiro.test)
  s <- do.call("c", lapply(l, "[[", 1))
  p <- do.call("c", lapply(l, "[[", 2))
  if (bonf == TRUE) {
    sig <- ifelse(p > alpha / length(l), "H0", "Ha")
  } else {
    sig <- ifelse(p > alpha, "H0", "Ha")
  }
  return(list(statistic= s,
              p.value= p,
              significance= sig,
              method= ifelse(bonf == TRUE, "Shapiro-Wilks test with Bonferroni Correction",
                             "Shapiro-Wilks test without Bonferroni Correction")))
}

shapiro_test_df(ks_vars_m)


#OR just look at a Normal Q-Q plot for each variable in the data frame

for (i in colnames(ks_vars_n))
     {
 
       normlpot <- qqnorm(ks_vars_n[,i])
}
    

```



WITHIN POPULATIONS
Test correlation of variables for each population separately, then match up the variables that are left.
Use KS test to compare distributions of variables between ENCOUNTERS in a population

```{r}
#Using the uncorrelated, continuous variables from the previous section, compare values between the encounters within the populations with the KS test again

#For pairwise KS tests between encounters

#Separate data by encounter






```

## Pairwise Mann-Whitney 
```{r}
MannWhitP <- pairwise.wilcox.test(pcdata.P_fmax$FREQMAX, pcdata.P_fmax$EncounterID, p.adjust.methods = "bonferroni")
MannWhitP

MannWhitP = MannWhitP$p.value

#Convert it to a full table
MannWhitP1 <- fullPTable(MannWhitP)
MannWhitP1
write.csv(MannWhitP1,'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_MannWhitTests.csv', row.names = F)
### Produce compact letter display. 
### Values sharing a letter ARE NOT significantly different. Groups 'ab' are not significantly different from groups 'a' or 'b'
library(multcompView)
multcompLetters(MannWhitP1,
                compare="<",
                threshold=0.05,  # p-value to use as significance threshold
                Letters=letters,
                reversed = FALSE)

```

```{r}
#Working with the uncorrelated data by population, subset it 

mw_M <- pcdata.Muncorr[, -c(2, 4:6)]
mw_M_results <- lapply(3:14, function(x) pairwise.wilcox.test(mw_M[[x]], mw_M$EncounterID, p.adjust.method = 'bonferroni'))
names(mw_M_results) <- names(mw_M)[3:14]
mw_M_results



mw_N <- pcdata.Nuncorr[, -c(2, 4:6)]
mw_N_results <- lapply(3:16, function(x) pairwise.wilcox.test(mw_N[[x]], mw_N$EncounterID, p.adjust.method = 'bonferroni'))
names(mw_N_results) <- names(mw_N)[3:16]
mw_N_results
```

out <- lapply(7:44, function(x) pairwise.wilcox.test(kwP_new[[x]], kwP_new$EncounterID, p.adjust.method = 'bonferroni'))
names(out) <- names(kwP_new)[7:44]
out
