---
title: "Multiple Comparison Tests"
subtitle: Chapter 1 - Pseudorca Whistle Analysis
output: html_document
---



#Kruskal-Wallis Tests
```{r}
library(dplyr)
library(data.table)

pcdata <- read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data/2017- 100 New Whistles\\ROCCA Towed Data\\BigRoccaFile_PMN_TowedAll_EDIT_20180914.csv')

pcdata.MN <- filter(pcdata, population == "MHI" | population == "NWHI")
pcdata.PM <- droplevels(filter(pcdata, population == "MHI" | population == "Pelagic"))
pcdata.PN <- droplevels(filter(pcdata, population == "Pelagic" | population == "NWHI"))


pcdata.P <- droplevels(filter(pcdata, population == "Pelagic"))      
pcdata.N <- droplevels(filter(pcdata, population == "NWHI"))      
pcdata.M <- droplevels(filter(pcdata, population == "MHI"))

```


##Normality test (controversial)
```{r}
normtest <- apply(pcdata[-c(1:6,54:61)],2, function(x) shapiro.test(x))
library(broom)
tidy  (normtest)
bart_pval <- do.call(rbind, lapply(bart, function(v){v$p.value})) 
bart_K2 <- do.call(rbind, lapply(bart, function(v){v$statistic})) #pull out only chi square values from list (makes matrix)
bart_df <- do.call(rbind, lapply(bart, function(v){v$parameter})) #pull out only degrees of freedom values from list (makes matrix)
bart_dat <- data.frame(names, bart_K2, bart_df, bart_pval, row.names = NULL)
```


##Bartlett's test to compare variances of populations
```{r}
bart <- apply(pcdata[-c(1:6,54:61)],2, function(x) bartlett.test(x,pcdata$population))

bart_pval <- do.call(rbind, lapply(bart, function(v){v$p.value})) 
bart_K2 <- do.call(rbind, lapply(bart, function(v){v$statistic})) #pull out only chi square values from list (makes matrix)
bart_df <- do.call(rbind, lapply(bart, function(v){v$parameter})) #pull out only degrees of freedom values from list (makes matrix)
bart_dat <- data.frame(names, bart_K2, bart_df, bart_pval, row.names = NULL)

#which variables are >= 0.05 (have similar variance)?
unique(bart_dat$names[bart_dat$bart_pval >= 0.05])

##variances are similar only for FREQBEGSWEEP    FREQBEGDWN      FREQENDUP       NUMSWEEPSUPFLAT    NUMSWEEPSFLATUP

```


##KW test WITHIN populations 
###May-Collado and Wartzok 2008
```{r}
#https://stackoverflow.com/questions/21271449/how-to-apply-the-wilcox-test-to-a-whole-dataframe-in-r

##FOR PELAGIC ENCOUNTERS
#Put column names of ROCCA variables in a column vector
names <- names(pcdata.P)[-c(1:6, 54:61)]

#This makes a list of all K-W test results for all variables
kwP <- apply(pcdata.P[-c(1:6,54:61)],2, function(x) kruskal.test(x,pcdata.P$EncounterID))

#I only want p-values < 0.05, pull out only pvalues from list (makes matrix)
kwP_pval <- do.call(rbind, lapply(kwP, function(v){v$p.value})) 
kwP_x2 <- do.call(rbind, lapply(kwP, function(v){v$statistic})) #pull out only chi square values from list (makes matrix)
kwP_df <- do.call(rbind, lapply(kwP, function(v){v$parameter})) #pull out only degrees of freedom values from list (makes matrix)
kwP_dat <- data.frame(names, kwP_x2, kwP_df, kwP_pval, row.names = NULL) #create dataframe 
colnames(kwP_dat)<-c("variable", "chi-squared", "df", "p.value")  #success!
#Add column with population name
kwP_dat$population <- 'Pelagic'
#filter out only variables with pvalues <= 0.05
kwP_signif <- droplevels(filter(kwP_dat, as.numeric(kwP_dat$p.value) <= 0.05) )
#namesP <- colnames(kwP_signif) #this is used to put new dataframe together later


#FOR NWHI ENCOUNTERS
kwN <- apply(pcdata.N[-c(1:6,54:61)],2, kruskal.test,pcdata.N$EncounterID)
kwN_pval <- do.call(rbind, lapply(kwN, function(v){v$p.value})) #pull out only pvalues from list (makes matrix)
kwN_x2 <- do.call(rbind, lapply(kwN, function(v){v$statistic})) #pull out only chi square values from list (makes matrix)
kwN_df <- do.call(rbind, lapply(kwN, function(v){v$parameter})) #pull out only degrees of freedom values from list (makes matrix)
kwN_dat <- data.frame(names, kwN_x2, kwN_df, kwN_pval, row.names = NULL) #create dataframe of pvalues from matrix
colnames(kwN_dat)<-c("variable", "chi-squared", "df", "p.value")  #success!
#Add column with population name
kwN_dat$population <- 'NWHI'
kwN_signif <- droplevels(filter(kwN_dat, as.numeric(kwN_dat$p.value) <= 0.05))
namesN <- colnames(kwN_signif)

#FOR MHI ENCOUNTERS 
kwM <- apply(pcdata.M[-c(1:6,54:61)],2, kruskal.test,pcdata.M$EncounterID)
kwM_pval <- do.call(rbind, lapply(kwM, function(v){v$p.value})) #pull out only values from list (makes matrix)
kwM_x2 <- do.call(rbind, lapply(kwM, function(v){v$statistic})) #pull out only chi square values from list (makes matrix)
kwM_df <- do.call(rbind, lapply(kwM, function(v){v$parameter})) #pull out only degrees of freedom values from list (makes matrix)
kwM_dat <- data.frame(names, kwM_x2, kwM_df, kwM_pval, row.names = NULL) #create dataframe of pvalues from matrix
colnames(kwM_dat)<-c("variable", "chi-squared", "df", "p.value")
#Add column with population name
kwM_dat$population <- 'MHI'
kwM_signif <- droplevels(filter(kwM_dat, as.numeric(kwM_dat$p.value) <= 0.05))
namesM <- colnames(kwM_signif)


#Combine KW tests for each population
kwAll <- rbind(kwP_dat, kwN_dat, kwM_dat)
kwAll <- kwAll[, c(5, 1:4)]
#Save all KW test for each population
write.csv(kwAll, 'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_KruskalWallis_All.csv')


kwAll_signif <- rbind(kwP_signif, kwN_signif, kwM_signif)
kwAll_signif <- kwAll_signif[, c(5, 1:4)]
#Save all KW test for each population
write.csv(kwAll_signif, 'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_KruskalWallis_AllSignif.csv')


```
##Subset data for only significant variables from KW test
```{r}
#Create dataframes with only variables that match the significant variables from the KW tests

#Within Pelagic
kwPvars <- kwP_signif[,1]
idx <- match(kwPvars, names(pcdata.P))  
kwP_new <- cbind(pcdata.P[,c(1:6)], pcdata.P[, idx])

#Within NWHI
kwNvars <- kwN_signif[,1]
idx <- match(kwNvars, names(pcdata.N))  
kwN_new <- cbind(pcdata.N[,c(1:6)], pcdata.N[, idx])

#Within MHI
kwMvars <- kwM_signif[,1]
idx <- match(kwMvars, names(pcdata.M))  
kwM_new <- cbind(pcdata.M[,c(1:6)], pcdata.M[, idx])


write.csv(kwP_new, 'C:\\Users\\Yvonne\\OneDrive\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles\\ROCCA Towed Data\\SIGNIF_P.csv', row.names = F)
write.csv(kwN_new, 'C:\\Users\\Yvonne\\OneDrive\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles\\ROCCA Towed Data\\SIGNIF_N.csv', row.names = F)
write.csv(kwM_new, 'C:\\Users\\Yvonne\\OneDrive\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles\\ROCCA Towed Data\\SIGNIF_M.csv', row.names = F)

```





##Dunns Test for Multiple Comparisons WITHIN population
```{r}
#Reference
#https://rcompanion.org/rcompanion/d_06.html
library(FSA)
library(rcompanion)

#For loop to get dunns test for all variables plus the letter list
#Not all variables are significantly different for some populations (NWHI & MHI), so comment out the dunn_letters portion to complete the Dunn's Tests for all variables. 
dunn_VARS <- kwP_new[,7:44] #pelagic
dunn_VARS <- kwN_new[,7:30] #NWHI
dunn_VARS <- kwM_new[,7:32] #MHI


#If a population has some variables that ARE NOT significantly different, an error in dunn_letters will occur
#Find all variables with pvalue <= 0.05 and create a new dataframe, dunn_VARS, to check out which pairs of encounters have signif differences.
idx2 = unique(dunn_resultsALL$variable[dunn_resultsALL$P.adj <= 0.05])
dunn_VARS <- pcdata.M[, idx]


dunn_lettersALL <- NULL
dunn_resultsALL <-NULL
for (v in colnames(dunn_VARS)){
  dunn_v <- dunnTest(dunn_VARS[, v] ~ EncounterID, data=kwM_new, method = 'bonferroni') #change data to  match population

  dunn_v = dunn_v$res
  dunn_v$variable <- v
  dunn_resultsALL <- rbind(dunn_resultsALL, dunn_v)

  dunn_letters <- cldList(comparison = dunn_v$Comparison,
                         p.value = dunn_v$P.adj,
                         threshold = 0.05)
  dunn_letters$variable <- v
  dunn_lettersALL <- rbind(dunn_lettersALL, dunn_letters)

}

write.csv(dunn_resultsALL, 'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_Dunns_mhi_signif.csv', row.names = F)

write.csv(dunn_lettersALL, 'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_DunnsLetters_mhi.csv', row.names = F)


###########old Dunns code without loop#############
#testing code using only freq max
pcdata.P_fmax <- pcdata.P[, c(1,3,4,7)]

boxplot(FREQMAX ~ EncounterID,
        data = pcdata.P_fmax,
        ylab="Maximum Frequency",
        xlab="Acoustic Encounter")
kruskal.test(FREQMAX ~ EncounterID, data = pcdata.P_fmax)

dunnP.bonf <- dunnTest(FREQMAX ~ EncounterID, data = pcdata.P_fmax, method="bonferroni")
dunnP.bonf

dunnP.bonf = dunnP.bonf$res
write.csv(dunnP.bonf,'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_DunnPvals.csv', row.names = F)

dunnlist <-cldList(comparison = dunnP.bonf$Comparison,
        p.value = dunnP.bonf$P.adj,
        threshold = 0.05)
write.csv(dunnlist,'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_DunnTests.csv', row.names = F)

### Dunn's Test for using all variables 
dunnP <- lapply(7:44, function(x) dunnTest(kwP_new[[x]], kwP_new$EncounterID, method = 'bonferroni'))
names(dunnP) <- names(kwP_new)[7:44]


dunnPdf <- sapply(dunnP, "[", c(2)) #list of dataframes
dunnPdf <- as.data.frame(dunnPdf)

write.csv(dunnPdf,'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_test.csv', row.names = F)

#list2env( setNames( test , paste('Pelagic', names(test), sep = '')), .GlobalEnv )


### Dunn's Test - NWHI
dunnN <- lapply(7:30, function(x) dunnTest(kwN_new[[x]], kwN_new$EncounterID, method = 'bonferroni'))
names(dunnN) <- names(kwN_new)[7:30]

dunnNdf <- sapply(dunnN, "[", c(2)) #list of dataframes
dunnNdf <- as.data.frame(dunnNdf)

write.csv(dunnNdf,'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_Dunns_nwhi.csv', row.names = F)

### Dunn's Test - MHI
dunnM <- lapply(7:32, function(x) dunnTest(kwM_new[[x]], kwM_new$EncounterID, method = 'bonferroni'))
names(dunnM) <- names(kwM_new)[7:32]

dunnMdf <- sapply(dunnM, "[", c(2)) #list of dataframes
dunnMdf <- as.data.frame(dunnMdf)

write.csv(dunnMdf,'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_Dunns_mhi.csv', row.names = F)




```


##Mann-Whitney U-tests for BETWEEN populations (only comparing 2 at a time)
```{r}

MannWhit_VARS <- pcdata[, c(7:53)]
#MannWhit_lettersALL <- NULL
MannWhit_resultsALL <-NULL
MannWhit_resultsALL2 <-NULL

for (k in colnames(MannWhit_VARS)){
  MannWhit_k <- pairwise.wilcox.test(MannWhit_VARS[, k], pcdata.MN$population, data=pcdata, p.adjust.method = 'bonferroni')
  # MannWhit_v = MannWhit_v$res
  # MannWhit_v$variable <- v
  MannWhit_k = MannWhit_k$p.value

#Convert it to a full table
MannWhit_k <- fullPTable(MannWhit_k)
  
MannWhit_k <- cbind(MannWhit_k, k)
  
  MannWhit_resultsALL2 <- rbind(MannWhit_resultsALL2, MannWhit_k)
  
  # MannWhit_letters <- cldList(comparison = MannWhit_v$Comparison,
  #                        p.value = MannWhit_v$P.adj,
  #                        threshold = 0.05)
  # MannWhit_letters$variable <- v
  # MannWhit_lettersALL <- rbind(MannWhit_lettersALL, MannWhit_letters)

}

write.csv(MannWhit_resultsALL, 'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_MannWhitney_All.csv', row.names = T)


```



##FREQQUARTER3 kruskal test
```{r}
#### FREQQUARTER3 ####

#MHI-NWHI-Pelagic
kruskal.test(FREQQUARTER3~population, pcdata)

#MHI-NWHI
kruskal.test(FREQQUARTER3~population, pcdata.MN)

#MHI-Pelagic
kruskal.test(FREQQUARTER3~population, pcdata.MP)

#Pelagic-NWHI
kruskal.test(FREQQUARTER3~population, pcdata.PN)
```




##DURATION kruskal test
```{r}
#### DURATION ####

#MHI-NWHI-Pelagic
kruskal.test(population ~ DURATION, pcdata)

#MHI-NWHI
kruskal.test(population ~ DURATION, pcdata.MN)

#MHI-Pelagic
kruskal.test(population ~ DURATION, pcdata.MP)

#Pelagic-NWHI
kruskal.test(population ~ DURATION, pcdata.PN)
```


## mann-Whitney U test with Bonferroni adjustment of pvalue (NOT using Holm method pvalue adjustment)
```{r}
out <- lapply(7:44, function(x) pairwise.wilcox.test(kwP_new[[x]], kwP_new$EncounterID, p.adjust.method = 'bonferroni'))
names(out) <- names(kwP_new)[7:44]
out

outtest <- list.group(out, p.value)

### Mann-Whitney Test - NWHI
out.N <- lapply(7:30, function(x) pairwise.wilcox.test(kwN_new[[x]], kwN_new$EncounterID, p.adjust.method = 'bonferroni'))
names(out.N) <- names(kwN_new)[7:30]
out.N


#matrix of only pvalues
outpval = sapply(out, function(x) {
    p <- x$p.value
    n <- outer(rownames(p), colnames(p), paste, sep='v')
    p <- as.vector(p)
    names(p) <- n
    p
})
outpval

write.csv(outpval,'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_MannWhitPvals.csv')

```




## Pairwise Mann-Whitney 
```{r}
MannWhitP <- pairwise.wilcox.test(pcdata.P_fmax$FREQMAX, pcdata.P_fmax$EncounterID, p.adjust.methods = "bonferroni")
MannWhitP

MannWhitP = MannWhitP$p.value

#Convert it to a full table
MannWhitP1 <- fullPTable(MannWhitP)
MannWhitP1
write.csv(MannWhitP1,'C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\KruskalWallisTests\\Pc_MannWhitTests.csv', row.names = F)
### Produce compact letter display. 
### Values sharing a letter ARE NOT significantly different. Groups 'ab' are not significantly different from groups 'a' or 'b'
library(multcompView)
multcompLetters(MannWhitP1,
                compare="<",
                threshold=0.05,  # p-value to use as significance threshold
                Letters=letters,
                reversed = FALSE)

```

