---
title: "R Notebook"
output: html_notebook
---

```{r warning=FALSE, message=FALSE}

library(randomForest)
library(caret)
library(corrplot)
library(parallel); 
library(doParallel);
library(gtools)
library(devtools) 
library(dplyr)
#library(ROCR)
library(Hmisc)


```


#Set up encounters for RF
```{r}
##Uncorrelated variables for all Random Forests
#pcdata<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\pcdata_20180809.csv')



#Laptop directory
pcdata_signif <- read.csv('C:\\Users\\Yvonne\\OneDrive\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles\\ROCCA Towed Data\\SIGNIF_PNM.csv')

#desktop directory
#



# Need even number of encounters for each population represented in the training data. 
PEL = 1:8                # designated groupID's assigned for each pelagic group 
NWHI  = 14:17   # designated groupID's assigned for each NWHI group, #20 was absorbed into #19 (June 2017)     
MHI = 31:34
                              # Oct 7: more combining of NWHI and MHI occurred. Performed in Pc_DatManipulation.rmd
#This is set up to be flexible to change later if needed
nrep  = 4           # multiplier for total number of groups
npel  = 1           # number of pelagic schools to group together
nnwhi = 1           # number of northwest HI schools to group together
nmhi  = 1           # number of main HI schools to group together
w = 150             # Number of whistles randomly pulled from each group 
ntest = 1           # number of groups to randomly pull for test dataset
n_samps = 5        #Zack: number of times you want to resample from your dataset

```


Meat of the Analysis

```{r warning=FALSE, message=FALSE}
#### Start ####
ptm<-proc.time()


results = matrix(0, nrow = 3, ncol = 3)
#output_ModelFit = matrix(0, nrow=3, ncol = 4)
output_ModelFit = NULL
realP = c()
realN = c()
realM = c()
#use this format of output_WhistleClass on desktop (R version 3.3.1)
output_WhistleClass = NULL
#use this format of output_WhistleClass on laptop (R version 3.5.1)
# output_WhistleClass = data.frame(Doubles=double(),
#                  Ints=integer(),
#                  Factors=factor(),
#                  Logicals=logical(),
#                  Characters=character(),
#                  stringsAsFactors=FALSE)
#output_EncounterClass=NULL
output_Results=NULL
output_TrainSetAll=NULL
output_TestSetAll = NULL
output_VotesAll=NULL
tunedAll=NULL
output_varIMPtotal = NULL
accuracies = c()

se <- function(x) sd(x)/sqrt(length(x))
# names.test=NULL
# names.train=NULL
#output_IndDWAll=NULL

#### OVERALL LOOP TO DERIVE MEANS AND STANDARD DEVS ####  
for(rep in 1:n_samps){ 
  
  #set.seed(rep) #YB placed the seed inside the loop bc it wasn't duplicating trees
  
  # Total groups included
  #This randomly samples 'nrep' groups from each population
  N_pel = as.numeric(sample(PEL, nrep*npel))
  N_nwhi  = as.numeric(sample(NWHI, nrep*nnwhi))
  N_mhi  = as.numeric(sample(MHI, nrep*nmhi))
  
  # Test data 
  #Randomly sample 1 group out of the selected groups for each population to be included in the test dataset later
  test_pel= as.integer(sample(N_pel, ntest))
  test_nwhi= as.integer(sample(N_nwhi, ntest))
  test_mhi= as.integer(sample(N_mhi, ntest))
  N_test = rbind(test_pel, test_nwhi, test_mhi)
  
  #Zack: I just combined the mhi, pelagic, and nwhi bit into one loop
  all = NULL
  for(i in c(N_pel,N_nwhi,N_mhi)){
    sub=droplevels(subset(pcdata_signif, pcdata_signif$group==i)) #selects all whistles from the randomly selected groups, drops empty levels too
    samp=sub[sample(nrow(sub), w),] #randomly samples w whistles from the 50 or more whistles
    all <-rbind(all, samp)
    
    }
  
#Correlation Step####
pcsub<-all[,7:length(pcdata_signif)]
#corrplot.mixed(cor(pcsub, method= "pearson"), upper="number", lower="circle")
tmp <- cor(pcsub)
tmp[upper.tri(tmp)] <- 0
diag(tmp) <- 0
pcsubcor <- pcsub[,!apply(tmp,2,function(x) any(x > 0.80 | x < -0.80))]

all <- cbind(all[, c(1:6)], pcsubcor)

  
  
  test<-droplevels(subset(all, group%in%N_test))
  train<-droplevels(subset(all, !(group%in%N_test)))
  #print(unique(train$EncounterID))
  #print(unique(test$EncounterID))
  
  # testp <- subset(test, test$population == 'pelagic')
  # testn <- subset(test, test$population == 'nwhi')

# ADDING } to test making the training and test data
#}
  
  
 


#### Tuning parameters ####
## Set up a dataframe where each row is a specific combination of mtry (# of variables selected at each split), n_tree (# of trees in  forest), possibly other stuff
mtry_vals = 5:8
ntree_vals = seq(500,5000,500) #from, to, by
tune_settings = expand.grid(mtry = mtry_vals, ntree = ntree_vals)
#tune_settings2 = expand.grid(mtry = mtry_vals, ntree = ntree_vals)

#### Run Optimization Sequence ####
## Initiate Cluster

cl = makeCluster(detectCores())
registerDoParallel(cl, cores = detectCores())


#### Formula ####
# for the RF model with 'population' as the response and time-frequency measurements 

## UNCORRELATED VARIABLES
formPc <- as.formula(paste("population ~ ", paste(names(all)[7:length(all)],collapse="+")))


## CORRELATED VARIABLES
#formPc <- as.formula(paste("population ~ ", paste(names(pcdata)[7:54],collapse="+"))) #Use variables to classify to population



##The foreach function is the loop function for parallel processing. 
##It uses %dopar% which partitions the loop among your cores, then combines each output using the .combine argument. 

x = foreach(i = 1:nrow(tune_settings), .combine = rbind) %dopar% { 
    library(randomForest)
    do_RF = randomForest(formPc, data = train, test = test, ntree = tune_settings$ntree[i], mtry = tune_settings$mtry[i], replace=F)
    print(do_RF$err.rate[tune_settings$ntree[i],])
  }

#so, for each of the parameter combinations, i.e., rows of the tune_settings df, the RF is run using those 
#settings (line 188) and the error rates for the 3 pop's and the oob are printed (line 189). 
#when all the rows are finished, the results are combined using the rbind function into one matrix, stored as x.

#NB: .combine can be anything, really, for example .combine = c for a vector, .combine = list for a list, etc.


#### Stop Cluster ####

stopCluster(cl)

## Combine settings and results ####

tune_settings = cbind(tune_settings, x)


## Optimal Settings: the settings combo
## that has the lowest OOB rate

tuned<-tune_settings[which.min(tune_settings$OOB),]
tunedAll <- rbind(tunedAll, tuned)

#### Run Random Forest ####
##Run Random Forest model using 'tuned' parameters for mtry and ntree
  fit <- randomForest(formPc, data=train, ntree=tuned[1,2], importance=TRUE, replace=F, mtry=tuned[1,1], norm.votes=F, confusion=T, keep.inbag=T, proximity=T, keep.forest=T, scale = F) 

#prediction for population of test data using RF model, 'fit', returns a list with aggregate and individual votes for each whistle  
  prediction <-predict(fit, newdata=test, predict.all = F, type='response') 
  prediction.vote <-predict(fit, newdata=test, predict.all = T, type='vote')
  
   test$rightPred <- as.character(prediction) == as.character(test$population) #shows T/F for individual whistles and whether they were classified correctly
  accuracy <- sum((test$rightPred)/nrow(test)) #sums all 'TRUE' and divides by total
  
  # results <- table(test$population, prediction)  # conf matrix of individual whistle classification
  
  #sums all 'TRUE' and divides by total
   accuracy <- sum((test$rightPred)/nrow(test)) 
   accuracies <- c(accuracies, accuracy) #correct classification
  ####RESULTS!!!#####  
  results.temp <- table(test$population, prediction)  # conf matrix of individual whistle counts
  
  #Summed results confusion matrix after each iteration
results = results + results.temp 
    
#For calculating means and standard dev for EACH ITERATION of confusion matrix
#For a single iteration, stores separate cells of class results for calculating mean and standard error
#YOU NEED THIS STEP 4/26/18, TRUST ME!
realP <- c(realP,results.temp[1,1])
#fakeP <- c(fakeP,results.temp[1,2])
realN <- c(realN,results.temp[2,2])
#fakeN <- c(fakeN,results.temp[2,1])
realM <- c(realM,results.temp[3,3])
#fakeM <- c(fakeM,results.temp[2,1])

#calculate % correct classification scores for each population 
pel.pel<-round((results.temp[1,1])/(results.temp[1,1]+results.temp[1,2]+results.temp[1,3])*100,2) #pelagic
nwhi.nwhi<-round((results.temp[2,2])/(results.temp[2,1]+results.temp[2,2]+results.temp[2,3])*100,2) #nwhi
mhi.mhi<-round((results.temp[3,3])/(results.temp[3,1]+results.temp[3,2]+results.temp[3,3])*100,2) #mhi
1-(sum(diag(results.temp))/nrow(test))


#calculate misclassification for pelagic
pel.nwhi<-round((results.temp[1,2])/(results.temp[1,1]+results.temp[1,2]+results.temp[1,3])*100,2)
pel.mhi<-round((results.temp[1,3])/(results.temp[1,1]+results.temp[1,2]+results.temp[1,3])*100,2)

#calculate misclassification for nwhi
nwhi.pel<-round((results.temp[2,1])/(results.temp[2,1]+results.temp[2,2]+results.temp[2,3])*100,2)
nwhi.mhi<-round((results.temp[2,3])/(results.temp[2,1]+results.temp[2,2]+results.temp[2,3])*100,2)

#calculate misclassification for mhi
mhi.pel<-round((results.temp[3,1])/(results.temp[3,1]+results.temp[3,2]+results.temp[3,3])*100,2) 
mhi.nwhi<-round((results.temp[3,2])/(results.temp[3,1]+results.temp[3,2]+results.temp[3,3])*100,2)

#Consolidate back into confusion matrix
pel.row<-c(pel.pel,pel.nwhi, pel.mhi)
nwhi.row<-c(nwhi.pel, nwhi.nwhi, nwhi.mhi)
mhi.row<-c(mhi.pel, mhi.nwhi, mhi.mhi)
class.title<-c('pelagic', 'nwhi', 'mhi')

conf.mat<-data.frame(rbind("Pelagic"=pel.row, "NWHI"=nwhi.row, "MHI"=mhi.row))
colnames(conf.mat)=c("Pelagic", "NWHI", "MHI")
conf.mat

#Set a threshold for the percentage of whistles that must be correctly classified to classify the population
true.pel = if(pel.pel>=50) 'Pelagic' else 'Ambiguous'
true.nwhi = if(nwhi.nwhi>=50) 'NWHI' else 'Ambiguous'
true.mhi = if(mhi.mhi>=50) 'MHI' else 'Ambiguous'
true.enc = data.frame(rbind(pel.pel, nwhi.nwhi, mhi.mhi))
#d = as.data.frame(unique(test$EncounterID))
#d = as.data.frame(d[c(1:3), ])
e=c()
e=as.data.frame(rbind(true.pel, true.nwhi, true.mhi))
e=cbind(e, true.enc)
colnames(e) = c("EncounterClass", "PctCorrect") #, "EncounterID")

#f=as.data.frame(cbind(prediction.vote$aggregate, prediction.vote$individual))

        
#### OUTPUTS ####

#1.
#Save each RF model's confusion matrix and OOB error
output_ModelFit <- rbind(output_ModelFit, as.table(fit$confusion), mean(fit$err.rate))

#2.
#Save results for each individual whistle, including overall predictions based on majority vote of forest and individual tree predictions
g=cbind(rep, "EncounterID"= test$EncounterID, "EncounterCount"=test$EncounterCount, "Prediction"=test$rightPred, "PredictedPop"=prediction, as.data.frame(cbind(prediction.vote$aggregate, prediction.vote$individual))) 

output_WhistleClass <- bind_rows(list(output_WhistleClass, g), .id ="id")

#3.
#table keeping track of individual confusion matrices for all iterations
output_Results<-rbind(output_Results, rep, results.temp, accuracy=signif(accuracy, digits = 3)) 

#4.
#df to show encounters used in each training set  
output_TrainSet <- data.frame(rep, "EncounterID" = sort(unique(train$EncounterID))) #, "encounter" = sort(unique(train$group)))
output_TrainSetAll <-rbind(output_TrainSetAll, output_TrainSet)

#5.
#df to show encounters used in each test set
output_TestSet <- data.frame(rep, "EncounterID" = unique(test$EncounterID), e) #, "encounter" = sort(unique(train$group)))
output_TestSetAll <-rbind(output_TestSetAll, output_TestSet)

#6.
  # Important Variables
  varIMPtemp <- as.data.frame(importance(fit))
  varIMP <- varIMPtemp[order(varIMPtemp$MeanDecreaseAccuracy), , drop =F]
  output_varIMPtotal <- rbind(output_varIMPtotal, "REP"=rep, varIMP)


#df to show the prediction of each individual whistle per encounter (THIS IS DOCUMENTED IN 'g')
# output_IndDW<-data.frame(rep, "EncounterID"=test$EncounterID, "PredictedPop"=prediction, "EncounterCount"=test$EncounterCount) 
# output_IndDWAll <- rbind(output_IndDWAll, output_IndDW)  #consolidates all individual whistle results for each iteration

}

correctP <- round((results[1,1]/(results[1,1]+results[1,2]+results[1,3])*100), 2)
correctN <- round((results[2,2]/(results[2,1]+results[2,2]+results[2,3])*100), 2)
correctM <- round((results[3,3]/(results[3,1]+results[3,2]+results[3,3])*100), 2)

output_CorrectClass <- rbind(results,cbind(correctP, correctN, correctM))


 proc.time() - ptm
 
``` 

#Calculate means and standard errors
```{r}
se <- function(x) sd(x)/sqrt(length(x))

meanRealP <- mean(realP)
meanRealN <- mean(realN)
meanRealM <- mean(realM)
seRealP <- se(realP)
seRealN <- se(realN)
seRealM <- se(realM)

 
date='20180824'
#Combine means and std errors into matrix
sink("C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018\\ConfMatrixPMN_means_20180824.txt", append=T)
PNMmeans <- cbind("PEL"= c(meanRealP), "NWHI" = c(meanRealN), "MHI"=c(meanRealM)) #means
PNMse <- cbind("PEL"= c(seRealP), "NWHI" = c(seRealN), "MHI"=c(seRealM)) #standard errors
PNMtotal <- rbind("Means" = PNMmeans, "Stand Err" = PNMse)

date
w
n_samps
PNMtotal
sink()



#For saving to laptop 
# date='20180818'
# #Combine means and std errors into matrix
# sink("C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\results\\20180818\\ConfMatrix_means_20180809.txt", append=T)
# PNMmeans <- cbind("PEL"= c(meanRealP), "NWHI" = c(meanRealN), "MHI"=c(meanRealM)) #means
# PNMse <- cbind("PEL"= c(seRealP), "NWHI" = c(seRealN), "MHI"=c(seRealM)) #standard errors
# PNMtotal <- rbind("Means" = PNMmeans, "Stand Err" = PNMse)
# 
# date
# w
# n_samps
# PNMtotal
# sink()
#  
``` 
 
 
###Write outputs to OneDrive
```{r}

#Save overall confusion matrix
write.table(output_CorrectClass, "C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018/PcResults_ConfusionMatrix_PMN_20180824.csv", append = T, row.names = T, col.names=T, sep = ",")

write.table(output_Results, "C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018/PcResults_TotalResults_PMN_20180824.csv", append = T, row.names = T, col.names=T, sep = ",")

write.table(output_WhistleClass, "C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018/PcResults_TotalWhistleClass_PMN_20180824.csv", append = T, row.names = F, col.names=T, sep = ",")

write.table(output_TrainSetAll, "C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018/PcResults_TrainSets_PMN_20180824.csv", append = T, col.names=T, sep = ",")

write.table(output_TestSetAll, "C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018/PcResults_TestResults_PMN_20180824.csv", append = T, col.names=T, sep = ",")

write.table(output_varIMPtotal, "C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018/PcResults_ImportantVars_PMN_20180824.csv", append = T, col.names=T, sep = ",")

write.table(tunedAll, "C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018/PcResults_TunedParameters_PMN_20180824.csv", append = T, col.names=T, sep = ",")

write.table(output_ModelFit, "C:\\Users\\yvers\\OneDrive\\PHD\\CHP1-FKW\\data\\results\\2018/PcResults_TotalModels_PMN_20180824.csv", append = T, col.names=F, sep = ",")

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
