---
title: "False Killer Whale Whistle Classification- Random Forest Analysis"
author: "Yvonne Barkley"
date: "March 21, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

This script creates random forests using whistle datasets (either old or new datasets)



 Load your libraries
```{r warning=FALSE, message=FALSE}
library(sqldf)
library(randomForest)
library(partykit)
library(caret)
library(corrplot)
library(parallel); 
library(doParallel);
library(gtools)
library(devtools) 
library(dplyr)
library(ROCR)
library(Hmisc)
library(tree)
```



If using the raw data file, run this section to modify the columns
```{r}
#Load ROCCA data (already replaced 'NA' and '999' with '0')
ROCCAdata<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/BIG ROCCA FILE raw.csv')

#Read old dataset
#pcdata<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\Pc_MasterListEdit-1016labs.csv')

#pcdata<-droplevels(subset(pcdata, Recorder != 'PMRF'))


#Add species names to last columns and save
colnames(ROCCAdata)[93] <- "Dc_Dd"
colnames(ROCCAdata)[94] <- "Gm"
colnames(ROCCAdata)[95] <- "Pc"
colnames(ROCCAdata)[96] <- "Sa"
colnames(ROCCAdata)[97] <- "Sb"
colnames(ROCCAdata)[98] <- "Sc"
colnames(ROCCAdata)[99] <- "Sl"
colnames(ROCCAdata)[100] <- "Tt"

#Reexamining whistles with outlying values of slope means
# test<- filter(ROCCAdata, FREQPOSSLOPEMEAN  > 90000)
# test2<- filter(ROCCAdata, FREQNEGSLOPEMEAN <  -90000)

#Remove unnecessary columns

pcdata <- ROCCAdata[, -c(1:3, 5:7, 11:13, 21, 22, 34:37, 68:92, 101)]
pcdata %>% mutate_if(is.factor, as.character) -> pcdata
pcdata %>% mutate_if(is.character, as.factor) -> pcdata


#i <- sapply(pcdata, is.factor)
#pcdata[i] <- lapply(pcdata[i], as.character)

###Add Population column and reorder
pcdata$population <-NA
levels(pcdata$population) <- c(levels(pcdata$population), "pel", "mhi", "nwhi")
pcdata[pcdata$group %in% c(26:30), "population"] <- "mhi"
pcdata[pcdata$group %in% c(14:17), "population"] <- "nwhi"
pcdata[pcdata$group %in% c(1:8), "population"] <- "pel"
pcdata$population <- as.factor(pcdata$population)
pcdata<- pcdata[, c(61, 1:60)]

pcdata.pm <- droplevels(subset(pcdata, population == 'pel' | population == 'mhi'))

#Removing MHI29 (cluster 4 enc)
#pcdata.pm <- subset(pcdata.pm, group != 29)

# Id_count <- pcdata[, c(1,3)] #Check it out

#Just looking at things
# pel6 <- filter(pcdata, EncounterID == 'PcPEL6')
# length(pcdata$EncounterID[which(pcdata$EncounterID == 'PcNWHI17')])
# pcNA <- pcdata[which(pcdata == 'NA')]
```

Correlation Stuff
Will remove correlated variables based on the r value.
```{r}
##Calculate Pearson's correlation coefficient for all variables. Remove variables -0.70 > r > 0.70.
pcsub<-pdata.pm[,7:53]
corrplot.mixed(cor(pcsub, method= "pearson"), upper="number", lower="circle")
tmp <- cor(pcsub)
tmp[upper.tri(tmp)] <- 0
diag(tmp) <- 0
pcsub <- pcsub[,!apply(tmp,2,function(x) any(x > 0.80 | x < -0.80))]

  pdata.pm <- cbind(pdata.pm[, c(1:4)], pcsub)
  corrplot.mixed(cor(pdata.pm[,5:30]), upper="circle", lower="number") # for a visual representation of the correlation
```


#Set up encounters for RF using new or old whistle data
```{r}
#new whistles
#pcdata.pm<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/BIG ROCCA FILE.csv')

#old whistles
pcdata.pm<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/BIG ROCCA OLD raw.csv')

pcdata.pm <- droplevels(subset(pcdata.pm, population == 'pel' | population == 'mhi'))

# Need even number of encounters for each population represented in the training data. 
PEL = 1:8                # designated groupID's assigned for each pelagic group 
#NWHI  = 14:17   # designated groupID's assigned for each NWHI group, #20 was absorbed into #19 (June 2017)     
MHI = c(26:34)    # designated groupID's assigned for each MHI group, #25 absorbed into #26 (June 2017) 
                              # Oct 7: more combining of NWHI and MHI occurred. Performed in Pc_DatManipulation.rmd

nrep  = 8           # multiplier for total number of groups
npel  = 1           # number of pelagic schools to group together
nnwhi = 1           # number of northwest HI schools to group together
nmhi  = 1           # number of main HI schools to group together
w = 45              # Number of whistles randomly pulled from each group 
ntest = 1           # number of groups to randomly pull for test dataset
n_samps = 5         #number of times you want to resample from your dataset


```


Meat of the Analysis

```{r message=FALSE, warning=FALSE}

ptm<-proc.time() 


#### Start ####
results.pm = matrix(0, nrow = 2, ncol = 2)
realP = c()
fakeP = c()
realM = c()
fakeM = c()
output_WhistleClass.pm =NULL
output_EncounterClass.pm=NULL
output_Results.pm=NULL
output_TrainSet.pmAll=NULL
output_TestSet.pmAll = NULL
output_Votes.pmAll=NULL
tuned.pmAll=NULL
output_varIMPtotal.pm = NULL


#fit.pm <- vector("list", n_samps)
#conf.mat.pm = matrix(0, nrow = 2, ncol = 2)
# names.test=NULL
# names.train=NULL
#output_IndDWAll=NULL

#### OVERALL LOOP TO DERIVE MEANS AND STANDARD DEVS ####  
for(rep in 1:n_samps){ 
 
  set.seed(rep) #YB placed the seed inside the loop bc it wasn't duplicating trees
  
  # Total groups included
  #This randomly samples 'nrep' groups from each population
  N_pel = as.numeric(sample(PEL, nrep*npel))
  N_mhi  = as.numeric(sample(MHI, nrep*nmhi))
  #N_nwhi  = as.numeric(sample(NWHI, nrep*nnwhi))
 
  
  # Test data 
  #Randomly sample 1 group out of the selected groups for each population to be included in the test dataset later
  test_pel= as.integer(sample(N_pel, ntest))
  test_mhi= as.integer(sample(N_mhi, ntest))
  #test_nwhi= as.integer(sample(N_nwhi, ntest))
  
  #Combines test groups into single vector
  N_test = rbind(test_pel, test_mhi)
  
  all = NULL
  for(i in c(N_pel,N_mhi)){
    sub=droplevels(subset(pcdata.pm, pcdata.pm$group==i)) #selects all whistles from the randomly selected groups, drops empty levels too
    samp=sub[sample(nrow(sub), w),] #randomly samples w whistles from the 50 or more whistles
    all <-rbind(all, samp)
    
    }
  
  test.pm<-droplevels(subset(all, group%in%N_test ))
  train.pm<-droplevels(subset(all, !(group%in%N_test)))
  #print(unique(train$EncounterID))
  #print(unique(test$EncounterID))
  
  # testp <- subset(test, test$population == 'pelagic')
  # testn <- subset(test, test$population == 'nwhi')

# ADDING } to test making the training and test data
#}
  
  
#ptm<-proc.time() 


#### Tuning parameters ####
## Set up a dataframe where each row is a specific combination of mtry (# of variables selected at each split), n_tree (# of trees in  forest), possibly other stuff
mtry_vals = 5:10 
ntree_vals = seq(501,3001,500) #from, to, by
#node_vals=1:3

tune_settings = expand.grid(mtry = mtry_vals, ntree = ntree_vals)#, nodesize = node_vals)
#tune_settings2 = expand.grid(mtry = mtry_vals, ntree = ntree_vals)

#### Run Optimization Sequence ####
## Initiate Cluster

cl = makeCluster(detectCores())
registerDoParallel(cl, cores = detectCores())


#### Formula ####
# for the RF model with 'population' as the response and time-frequency measurements 

## UNCORRELATED VARIABLES
#formPc <- as.formula(paste("population ~ ", paste(names(pcdata.pm)[5:30],collapse="+")))

## CORRELATED VARIABLES
formPc <- as.formula(paste("population ~ ", paste(names(pcdata.pm)[7:54],collapse="+"))) #Use variables to classify to population



##The foreach function is the loop function for parallel processing. 
##It uses %dopar% which partitions the loop among your cores, then combines each output using the .combine argument. 

x = foreach(i = 1:nrow(tune_settings), .combine = rbind) %dopar% { 
    library(randomForest)
    do_RF = randomForest(formPc, data = train.pm, test = test.pm, ntree = tune_settings$ntree[i], mtry = tune_settings$mtry[i])#, nodesize = tune_settings$nodesize[i])
    print(do_RF$err.rate[tune_settings$ntree[i],])
  }

#so, for each of the parameter combinations, i.e., rows of the tune_settings df, the RF is run using those 
#settings (line 188) and the error rates for the 3 pop's and the oob are printed (line 189). 
#when all the rows are finished, the results are combined using the rbind function into one matrix, stored as x.

#NB: .combine can be anything, really, for example .combine = c for a vector, .combine = list for a list, etc.


#### Stop Cluster ####

stopCluster(cl)

## Combine settings and results ####

tune_settings = cbind(tune_settings, x)


## Optimal Settings: the settings combo
## that has the lowest OOB rate

tuned.pm<-tune_settings[which.min(tune_settings$OOB),]
tuned.pmAll <- rbind(tuned.pmAll, tuned.pm)

#### Run Random Forest ####
##Run Random Forest model using 'tuned' parameters for mtry and ntree
  fit.pm <- randomForest(formPc, data=train.pm, ntree=tuned.pm[1,2], importance=TRUE, replace=T, mtry=tuned.pm[1,1], norm.votes=F, confusion=T, keep.inbag=T, proximity=T, keep.forest=T, scale = F) 

#prediction for population of test data using RF model, 'fit', returns a list with aggregate and individual votes for each whistle  
  prediction.pm <-predict(fit.pm, newdata=test.pm, predict.all = F, type='response', nodes = T) 
  prediction.vote <-predict(fit.pm, newdata=test.pm, predict.all = T, type='vote')

   #shows T/F for individual whistles and whether they were classified correctly  
   test.pm$rightPred <- as.character(prediction.pm) == as.character(test.pm$population) 
  
   #sums all 'TRUE' and divides by total
   accuracy.pm <- sum((test.pm$rightPred)/nrow(test.pm)) 
  
  ####RESULTS!!!#####  
  results.pmtemp <- table(test.pm$population, prediction.pm)  # conf matrix of individual whistle counts
  
## Get percentages of whistles correctly and incorrectly classified   
  #calculate % accuracies of diagonal by population from confusion matrix of percentages
pel.pel<-round((results.pmtemp[1,1])/(results.pmtemp[1,1]+results.pmtemp[1,2])*100,2) #pelagic
mhi.mhi<-round((results.pmtemp[2,2])/(results.pmtemp[2,1]+results.pmtemp[2,2])*100,2) #mhi 
  #calculate misclassification for pelagic
pel.mhi<-round((results.pmtemp[1,2])/(results.pmtemp[1,1]+results.pmtemp[1,2])*100,2)
mhi.pel<-round((results.pmtemp[2,1])/(results.pmtemp[2,1]+results.pmtemp[2,2])*100,2) 
  

#Calculate means and standard dev for EACH ITERATION of confusion matrix
#For a single iteration, stores separate cells of results for calculating mean and standard error
#YOU NEED THIS STEP 4/26/18, TRUST ME!
realP <- c(realP,results.pmtemp[1,1])
fakeP <- c(fakeP,results.pmtemp[1,2])
realM <- c(realM,results.pmtemp[2,2])
fakeM <- c(fakeM,results.pmtemp[2,1])


#Consolidate back into confusion matrix
pel.row<-c(pel.pel, pel.mhi)
#nwhi.row<-c(nwhi.pel, nwhi.nwhi, nwhi.mhi)
mhi.row<-c(mhi.pel, mhi.mhi)
#class.title<-c('pelagic', 'mhi')
conf.mat.pm <- rbind(pel.row, mhi.row) #(% of whistles)
#conf.mat.pm = conf.mat.pm + conf.mat.pmtemp
#conf.mat.pm<-data.frame(rbind("Pelagic"=pel.row, "MHI"=mhi.row))


colnames(conf.mat.pm)=c("Pelagic", "MHI")
#conf.mat.pm

true.pel = if(pel.pel>60) 'Pelagic' else 'Ambiguous'
#true.nwhi = if(nwhi.nwhi>50) 'NWHI' else 'Ambiguous'
true.mhi = if(mhi.mhi>60) 'MHI' else 'Ambiguous'
true.enc = data.frame(rbind(pel.pel, mhi.mhi))
#d = as.data.frame(unique(test$EncounterID))
#d = as.data.frame(d[c(1:3), ])
e.pm=c()
e.pm=as.data.frame(rbind(true.pel, true.mhi))
e.pm=cbind(e.pm, true.enc)
colnames(e.pm) = c("EncounterClass", "% Correct") #, "EncounterID")

#Summed results confusion matrix after each iteration
results.pm = results.pm + results.pmtemp 


        
#### OUTPUTS ####

g.pm=cbind(rep, "EncounterID"= test.pm$EncounterID, "EncounterCount"=test.pm$EncounterCount, "Prediction"=test.pm$rightPred, "PredictedPop"=prediction.pm, as.data.frame(cbind(prediction.vote$aggregate, prediction.vote$individual))) 

#Most useful
#Output shows encounters in each test set, encounter prediction, and % correctly classified
output_TestSet.pm <- data.frame(rep, "EncounterID" = sort(unique(test.pm$EncounterID)), e.pm) #, "encounter" = sort(unique(train$group)))
output_TestSet.pmAll <-rbind(output_TestSet.pmAll, output_TestSet.pm)

#Most comprehensive information about overall predictions, proportional classification of trees, and tree predictions for each whistle
output_WhistleClass.pm <- smartbind(output_WhistleClass.pm, g.pm)

#Includes confusion matrix for each iteration (in number of whistles), the cumulative results.pm are already saved and converted to confusion matrix outside of loop below
output_Results.pm<-rbind(output_Results.pm, rep, results.pmtemp, accuracy.pm=signif(accuracy.pm, digits = 3)) 

#Output of encounters used in each training set  
output_TrainSet.pm <- data.frame(rep, "EncounterID" = sort(unique(train.pm$EncounterID))) #, "encounter" = sort(unique(train$group)))
output_TrainSet.pmAll <-rbind(output_TrainSet.pmAll, output_TrainSet.pm)

# Important Variables
  varIMPtemp.pm <- as.data.frame(importance(fit.pm))
  varIMP.pm <- varIMPtemp.pm[order(varIMPtemp.pm$MeanDecreaseGini), , drop =F]
  output_varIMPtotal.pm <- rbind(output_varIMPtotal.pm, "REP"=rep, varIMP.pm)


#rough table keeping track of confusion matrices for all iterations
#output_ConfMat.pm<-rbind(output_ConfMat.pm, rep, conf.mat.pm, accuracy.pm=signif(accuracy.pm, digits = 3))

#Unnecessary, redundant with output_TestSet.pm
#output_EncounterClass.pm <- rbind(output_EncounterClass.pm, e.pm)


#df to show the prediction of each individual whistle per encounter (THIS IS DOCUMENTED IN 'g')
# output_IndDW<-data.frame(rep, "EncounterID"=test$EncounterID, "PredictedPop"=prediction, "EncounterCount"=test$EncounterCount) 
# output_IndDWAll <- rbind(output_IndDWAll, output_IndDW)  #consolidates all individual whistle results for each iteration

}

####Mean & StDev ####
#After all forests are fit, take the cumulative confusion matrix and run some stats
#Calculates overall results for each iteration, summing up all whistle classifications
pel.diag<-round((results.pm[1,1])/(results.pm[1,1]+results.pm[1,2])*100,2) #pelagic
mhi.diag<-round((results.pm[2,2])/(results.pm[2,1]+results.pm[2,2])*100,2) #mhi
#calculate misclassification for pelagic and mhi
pel.mis<-round((results.pm[1,2])/(results.pm[1,1]+results.pm[1,2])*100,2)
mhi.mis<-round((results.pm[2,1])/(results.pm[2,1]+results.pm[2,2])*100,2)

pel.tot<-c(pel.diag, pel.mis)
mhi.tot<-c(mhi.mis, mhi.diag)
Conf.Mat.pm <- rbind(pel.tot, mhi.tot)
colnames(Conf.Mat.pm)=c("Pelagic", "MHI")

#Takes mean and standard dev of cumulative whistles correctly and incorrectly classified
se <- function(x) sd(x)/sqrt(length(x))

meanRealP <- mean(realP)
meanFakeP <- mean(fakeP)
meanRealM <- mean(realM)
meanFakeM <- mean(fakeM)
# seRealP <- se(realP)
# seFakeP <- se(fakeP)
# seRealM <- se(realM)
# seFakeM <- se(fakeM)
sdRealP <- sd(realP)
sdFakeP <- sd(fakeP)
sdRealM <- sd(realM)
sdFakeM <- sd(fakeM)

date='20180501 old'
#Combine means and std errors into matrix
sink("ConfMatrix_means_20180501.txt", append=T)
PMmeans <- rbind("PEL"= c(meanRealP, meanFakeP), "MHI"=c(meanRealM, meanFakeM))
colnames(PMmeans) <- c("PEL", "MHI")
date
w
n_samps
PMmeans
sink()

sink("ConfMatrix_stddevs_20180501.txt", append=T)
PMsds <- rbind("PEL"= c(sdRealP, sdFakeP), "MHI"=c(sdRealM, sdFakeM))
colnames(PMsds) <- c("PEL", "MHI")
date
w
n_samps
PMsds
sink()


proc.time() - ptm

```


###Write NEW WHISTLE DATA outputs to file
```{r}
write.table(output_EncounterClass.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TotalEncounterClass_newPvM.csv", append = T, row.names = T, col.names=T, sep = ",")

write.table(output_WhistleClass.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TotalWhistleClass_newPvM.csv", append = T, row.names = F, col.names=T, sep = ",")

#write.table(output_ConfMat.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TotalConfMatrix_newPvM.csv", append = T, col.names=F, sep = ",")

write.table(output_TrainSet.pmAll, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TrainSets_newPvM.csv", append = T, col.names=T, sep = ",")

write.table(output_TestSet.pmAll, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TestResults_newPvM.csv", append = T, col.names=T, sep = ",")

write.table(varIMPtotal.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_ImportantVars_newPvM.csv", append = T, col.names=T, sep = ",")
```
###Write OLD WHISTLE DATA outputs to file
```{r}
write.table(output_EncounterClass.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TotalEncounterClass_oldPvM.csv", append = T, row.names = T, col.names=T, sep = ",")

write.table(output_WhistleClass.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TotalWhistleClass_oldPvM.csv", append = T, row.names = F, col.names=T, sep = ",")

#write.table(output_ConfMat.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TotalConfMatrix_newPvM.csv", append = T, col.names=F, sep = ",")

write.table(output_TrainSet.pmAll, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TrainSets_oldPvM.csv", append = T, col.names=T, sep = ",")

write.table(output_TestSet.pmAll, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_TestResults_oldPvM.csv", append = T, col.names=T, sep = ",")

write.table(output_varIMPtotal.pm, "C://Users//Yvonne//Documents//PHD//CHP1-FKW//data//results//PcResults_ImportantVars_oldPvM.csv", append = T, col.names=T, sep = ",")
```


##Testing HARP data with new whistles
```{r warning=FALSE, message=FALSE}

train.pm<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/BIG ROCCA FILE.csv')
train.pm <- droplevels(filter(train.pm, population == 'mhi' | population == 'pel'))

test.pm<-read.csv('C:\\Users\\Yvonne\\Documents\\PHD\\CHP1-FKW\\data\\2017- 100 New Whistles/BIG ROCCA harp only.csv')

# Need even number of encounters for each population represented in the training data. 
PEL = 1:8                # designated groupID's assigned for each pelagic group 
NWHI  = 14:17   # designated groupID's assigned for each NWHI group, #20 was absorbed into #19 (June 2017)     
MHI = c(26:30)    # designated groupID's assigned for each MHI group, #25 absorbed into #26 (June 2017) 
                              # Oct 7: more combining of NWHI and MHI occurred. Performed in Pc_DatManipulation.rmd
#This is set up to be flexible to change later if needed
nrep  = 8           # multiplier for total number of groups
npel  = 1           # number of pelagic schools to group together
nnwhi = 1           # number of northwest HI schools to group together
nmhi  = 1           # number of main HI schools to group together
w = 45              # Number of whistles randomly pulled from each group 
ntest = 1           # number of groups to randomly pull for test dataset
n_samps = 2        #Zack: number of times you want to resample from your dataset


```


#Classify HARP data only
Meat of the Analysis
```{r message=FALSE, warning=FALSE}

ptm<-proc.time() 


#### Start ####
results.pm = matrix(0, nrow = 1, ncol = 2)
realP = c()
fakeP = c()
realM = c()
fakeM = c()
output_WhistleClass.pm =NULL
output_EncounterClass.pm=NULL
output_RawResults.pm=NULL
output_TrainSet.pmAll=NULL
output_TestSet.pmAll = NULL
output_Votes.pmAll=NULL
tuned.pmAll=NULL
varIMPtotal.pm = NULL


#fit.pm <- vector("list", n_samps)
#conf.mat.pm = matrix(0, nrow = 2, ncol = 2)
# names.test=NULL
# names.train=NULL
#output_IndDWAll=NULL

#### OVERALL LOOP TO DERIVE MEANS AND STANDARD DEVS ####  
 for(rep in 1:n_samps){ 
#  
#   set.seed(rep) #YB placed the seed inside the loop bc it wasn't duplicating trees
#   
#   # Total groups included
#   #This randomly samples 'nrep' groups from each population
#   N_pel = as.numeric(sample(PEL, nrep*npel))
#   N_mhi  = as.numeric(sample(MHI, nrep*nmhi))
#   #N_nwhi  = as.numeric(sample(NWHI, nrep*nnwhi))
#  
#   
#   # Test data 
#   #Randomly sample 1 group out of the selected groups for each population to be included in the test dataset later
#   test_pel= as.integer(sample(N_pel, ntest))
#   test_mhi= as.integer(sample(N_mhi, ntest))
#   #test_nwhi= as.integer(sample(N_nwhi, ntest))
#   
#   #Combines test groups into single vector
#   N_test = rbind(test_pel, test_mhi)
#   
#   all = NULL
#   for(i in c(N_pel,N_mhi)){
#     sub=droplevels(subset(pcdata.pm, pcdata.pm$group==i)) #selects all whistles from the randomly selected groups, drops empty levels too
#     samp=sub[sample(nrow(sub), w),] #randomly samples w whistles from the 50 or more whistles
#     all <-rbind(all, samp)
#     
#     }
#   
#   test.pm<-droplevels(subset(all, group%in%N_test ))
#   train.pm<-droplevels(subset(all, !(group%in%N_test)))
  #print(unique(train$EncounterID))
  #print(unique(test$EncounterID))
  
  # testp <- subset(test, test$population == 'pelagic')
  # testn <- subset(test, test$population == 'nwhi')

# ADDING } to test making the training and test data
#}
  
  
#ptm<-proc.time() 


#### Tuning parameters ####
## Set up a dataframe where each row is a specific combination of mtry (# of variables selected at each split), n_tree (# of trees in  forest), possibly other stuff
mtry_vals = 5:10 
ntree_vals = seq(501,3001,50) #from, to, by
#node_vals=1:3

tune_settings = expand.grid(mtry = mtry_vals, ntree = ntree_vals)#, nodesize = node_vals)
#tune_settings2 = expand.grid(mtry = mtry_vals, ntree = ntree_vals)

#### Run Optimization Sequence ####
## Initiate Cluster

cl = makeCluster(detectCores())
registerDoParallel(cl, cores = detectCores())


#### Formula ####
# for the RF model with 'population' as the response and time-frequency measurements 

## UNCORRELATED VARIABLES
#formPc <- as.formula(paste("population ~ ", paste(names(pcdata.pm)[5:30],collapse="+")))

## CORRELATED VARIABLES
formPc <- as.formula(paste("population ~ ", paste(names(pcdata.pm)[7:54],collapse="+"))) #Use variables to classify to population



##The foreach function is the loop function for parallel processing. 
##It uses %dopar% which partitions the loop among your cores, then combines each output using the .combine argument. 

x = foreach(i = 1:nrow(tune_settings), .combine = rbind) %dopar% { 
    library(randomForest)
    do_RF = randomForest(formPc, data = train.pm, test = test.pm, ntree = tune_settings$ntree[i], mtry = tune_settings$mtry[i])#, nodesize = tune_settings$nodesize[i])
    print(do_RF$err.rate[tune_settings$ntree[i],])
  }

#so, for each of the parameter combinations, i.e., rows of the tune_settings df, the RF is run using those 
#settings (line 188) and the error rates for the 3 pop's and the oob are printed (line 189). 
#when all the rows are finished, the results are combined using the rbind function into one matrix, stored as x.

#NB: .combine can be anything, really, for example .combine = c for a vector, .combine = list for a list, etc.


#### Stop Cluster ####

stopCluster(cl)

## Combine settings and results ####

tune_settings = cbind(tune_settings, x)


## Optimal Settings: the settings combo
## that has the lowest OOB rate

tuned.pm<-tune_settings[which.min(tune_settings$OOB),]
tuned.pmAll <- rbind(tuned.pmAll, tuned.pm)

#### Run Random Forest ####
##Run Random Forest model using 'tuned' parameters for mtry and ntree
  fit.pm <- randomForest(formPc, data=train.pm, ntree=tuned.pm[1,2], importance=TRUE, replace=T, mtry=tuned.pm[1,1], norm.votes=F, confusion=T, keep.inbag=T, proximity=T, keep.forest=T, scale = F) 

#prediction for population of test data using RF model, 'fit', returns a list with aggregate and individual votes for each whistle  
  prediction.pm <-predict(fit.pm, newdata=test.pm, predict.all = F, type='response', nodes = T) 
  prediction.vote <-predict(fit.pm, newdata=test.pm, predict.all = T, type='vote')

   #shows T/F for individual whistles and whether they were classified correctly  
   test.pm$rightPred <- as.character(prediction.pm) == as.character(test.pm$population) 
  
   #sums all 'TRUE' and divides by total
   accuracy.pm <- sum((test.pm$rightPred)/nrow(test.pm)) 
  
#### Accuracy for each EncounterID in test set####  
   encID = unique(test.pm$EncounterID)
   accID.pm = NULL
   encCLASS = NULL
   for (id in encID) {
     subID <- subset(test.pm, EncounterID == id)
     accID.pmTemp <- sum((subID$rightPred)/nrow(subID)) 
     encCLASStemp = if(accID.pmTemp>=.60) 'MHI' else 'Ambiguous'
     encCLASS = rbind.data.frame(encCLASS, encCLASStemp)
     accID.pm <- rbind.data.frame(accID.pm, accID.pmTemp)
     
   }
   accID.pmALL <- as.data.frame(cbind(rep, encID, accID.pm,  encCLASS))
   colnames(accID.pmALL) = c("Rep", "EncounterID", "Accuracy", "Class")
   
  ####RESULTS!!!#####  
  results.pmtemp <- table(test.pm$population, prediction.pm)  # conf matrix of individual whistle classification
  
  #calculate % accuracies of diagonal by population from confusion matrix of percentages
#pel.pel<-round((results.pmtemp[1,1])/(results.pmtemp[1,1]+results.pmtemp[1,2])*100,2) #pelagic
mhi.mhi<-round((results.pmtemp[1,1])/(results.pmtemp[1,1]+results.pmtemp[1,2])*100,2) #mhi 
  #calculate misclassification for pelagic
# pel.mhi<-round((results.pmtemp[1,2])/(results.pmtemp[1,1]+results.pmtemp[1,2])*100,2)
# mhi.pel<-round((results.pmtemp[2,1])/(results.pmtemp[2,1]+results.pmtemp[2,2])*100,2) 
  

  # Important Variables
  varIMPtemp.pm <- as.data.frame(importance(fit.pm))
  varIMP.pm <- varIMPtemp.pm[order(varIMPtemp.pm$MeanDecreaseGini), , drop =F]
  varIMPtotal.pm <- rbind(varIMPtotal.pm, "REP"=rep, varIMP.pm)

  
   
#Calculate means and standard dev for each iteration of confusion matrix
#se=function(x) sd(x)/sqrt(length(x))
#for a single iteration, stores separate cells of results for calculating mean and standard error
se <- function(x) sd(x)/sqrt(length(x))
# realP <- c(realP,results.pmtemp[1,1])
# fakeP <- c(fakeP,results.pmtemp[1,2])
realM <- c(realM,results.pmtemp[1,2])
fakeM <- c(fakeM,results.pmtemp[1,2])


#Consolidate back into confusion matrix
# pel.row<-c(pel.pel, pel.mhi)
# #nwhi.row<-c(nwhi.pel, nwhi.nwhi, nwhi.mhi)
# mhi.row<-c(mhi.pel, mhi.mhi)
# #class.title<-c('pelagic', 'mhi')
# conf.mat.pm <- rbind(pel.row, mhi.row)
# #conf.mat.pm = conf.mat.pm + conf.mat.pmtemp
# #conf.mat.pm<-data.frame(rbind("Pelagic"=pel.row, "MHI"=mhi.row))
# 
# 
# colnames(conf.mat.pm)=c("Pelagic", "MHI")
#conf.mat.pm

#true.pel = if(pel.pel>60) 'Pelagic' else 'Ambiguous'
#true.nwhi = if(nwhi.nwhi>50) 'NWHI' else 'Ambiguous'
true.mhi = if(mhi.mhi>60) 'MHI' else 'Ambiguous'
true.enc = data.frame(mhi.mhi)
#d = as.data.frame(unique(test$EncounterID))
#d = as.data.frame(d[c(1:3), ])
e.pm=c()
e.pm=as.data.frame(true.mhi)
e.pm=cbind(e.pm, true.enc)
colnames(e.pm) = c("EncounterClass", "% Correct") #, "EncounterID")

#Summed results confusion matrix after each iteration
results.pm = results.pm + results.pmtemp 


        
#### OUTPUTS ####
# BIG dataframe with RF results for each encounter, including the % of votes, individ tree votes
g.pm=cbind(rep, "EncounterID"= test.pm$EncounterID, "EncounterCount"=test.pm$EncounterCount, "Prediction"=test.pm$rightPred, "PredictedPop"=prediction.pm, as.data.frame(cbind(prediction.vote$aggregate, prediction.vote$individual))) 

output_WhistleClass.pm <- smartbind(output_WhistleClass.pm, g.pm)

output_EncounterClass.pm <- rbind(output_EncounterClass.pm, e.pm)

output_RawResults.pm<-rbind(output_RawResults.pm, rep, results.pm, accuracy.pm=signif(accuracy.pm, digits = 3)) #rough table keeping track of confusion matrices for all iterations
output_ConfMat.pm<-rbind(output_ConfMat.pm, rep, conf.mat.pm, accuracy.pm=signif(accuracy.pm, digits = 3))
#df to show encounters used in each training set  
output_TrainSet.pm <- data.frame(rep, "EncounterID" = sort(unique(train.pm$EncounterID))) #, "encounter" = sort(unique(train$group)))
output_TrainSet.pmAll <-rbind(output_TrainSet.pmAll, output_TrainSet.pm)

#df to show encounters used in each test set
output_TestSet.pm <- data.frame(rep, "EncounterID" = sort(unique(test.pm$EncounterID)), e.pm) #, "encounter" = sort(unique(train$group)))
output_TestSet.pmAll <-rbind(output_TestSet.pmAll, output_TestSet.pm)

#df to show the prediction of each individual whistle per encounter (THIS IS DOCUMENTED IN 'g')
# output_IndDW<-data.frame(rep, "EncounterID"=test$EncounterID, "PredictedPop"=prediction, "EncounterCount"=test$EncounterCount) 
# output_IndDWAll <- rbind(output_IndDWAll, output_IndDW)  #consolidates all individual whistle results for each iteration

}


#Calculates overall results for each iteration, summing up all whistle classifications
pel.diag<-round((results.pm[1,1])/(results.pm[1,1]+results.pm[1,2])*100,2) #pelagic
mhi.diag<-round((results.pm[2,2])/(results.pm[2,1]+results.pm[2,2])*100,2) #mhi
#calculate misclassification for pelagic and mhi
pel.mis<-round((results.pm[1,2])/(results.pm[1,1]+results.pm[1,2])*100,2)
mhi.mis<-round((results.pm[2,1])/(results.pm[2,1]+results.pm[2,2])*100,2)

pel.tot<-c(pel.diag, pel.mis)
mhi.tot<-c(mhi.mis, mhi.diag)
Conf.Mat.pm <- rbind(pel.tot, mhi.tot)
colnames(Conf.Mat.pm)=c("Pelagic", "MHI")


meanRealP <- mean(realP)
meanFakeP <- mean(fakeP)
meanRealM <- mean(realM)
meanFakeM <- mean(fakeM)
# seRealP <- se(realP)
# seFakeP <- se(fakeP)
# seRealM <- se(realM)
# seFakeM <- se(fakeM)
sdRealP <- sd(realP)
sdFakeP <- sd(fakeP)
sdRealM <- sd(realM)
sdFakeM <- sd(fakeM)

#Combine means and std errors into matrix
sink("ConfMatrix_means_20180409.txt", append=T)
PMmeans <- rbind("PEL"= c(meanRealP, meanFakeP), "MHI"=c(meanFakeM,meanRealM))
colnames(PMmeans) <- c("PEL", "MHI")
w
n_samps
PMmeans
sink()

sink("ConfMatrix_stddevs_20180409.txt", append=T)
PMsds <- rbind("PEL"= c(sdRealP, sdFakeP), "MHI"=c(sdFakeM,sdRealM))
colnames(PMsds) <- c("PEL", "MHI")
w
n_samps
PMsds
sink()


proc.time() - ptm

```



```


```{r}

  N_pel = as.numeric(sample(PEL, nrep*npel))
  N_mhi  = as.numeric(sample(MHI, nrep*nmhi))


#Subsetting 50dw from each encounter since some have >50 dw.
  
pw_pm = NULL

for(i in c(N_mhi)){
# for(i in c(N_pel)){   
 sub=droplevels(subset(pcdata, pcdata$group==i)) #selects all whistles from the randomly selected groups, drops empty levels too
    samp=sub[sample(nrow(sub), w),] #randomly samples w whistles from the 50 or more whistles
    pw_pm <-rbind(pw_pm, samp)  #this will stay the same
    
  }



P = subset(pcdata, pcdata$population=='pelagic')
M = NULL
for(i in c(MHI)){
 sub=droplevels(subset(pcdata, pcdata$group==i)) #selects all whistles from the randomly selected groups, drops empty levels too
    samp=sub[sample(nrow(sub), w),] #randomly samples w whistles from the 50 or more whistles
    M <-rbind(M, samp)  #this will stay the same
    
  }


```